---
title: Late-Night Gaming, Sleep and Wellbeing
subtitle: The Pri(n)ce of Playing Past Bedtime
running-head: Late-Night Gaming
author:
  - name: Tamás A. Földes
    email: tamas.foldes@oii.ox.ac.uk
    orcid: 0000-0002-0623-9149
    url: https://example.com
    corresponding: true
    affiliation:
      - ref: 1
  - name: Second Author
    affiliation:
      - ref: 2
affiliations:
  - id: 1
    name: First University
    department: Department Name
  - id: 2
    name: Second University
abstract: |
  Your abstract goes here. This will appear prominently at the top of your document.
categories: [keyword1, keyword2, keyword3]
authornote: |
  Author notes go here.
date: last-modified
bibliography: bibliography.bib
prefer-html: true
execute:
  echo: false
  warning: false
  message: false
  cache: true
  freeze: auto   # Use `quarto render --cache-refresh` to force re-execution
params:
  refit_panel: false
  refit_diary: false
  refit_diary_imputed: false
  refit_appendix: false
  rerun_preprocessing: false
format:
  html:
    toc: true
    toc-location: left
    theme: cosmo
    code-fold: true
    html-math-method: katex
    embed-resources: true
  preprint-typst:
    wordcount: true
    citeproc: true
    # Common options - uncomment to use
    theme: jou                   # Journal theme (2-column)
    # line-number: true            # Line numbering
    # fontsize: 11pt               # Font size
    # leading: 0.6em               # Line spacing
    # first-line-indent: 1.8em     # Paragraph indent
---

# Introduction

Concerns have been raised about the potential negative impacts of video gaming on sleep and overall wellbeing, particularly for adolescents and young adults and especially when gaming occurs late at night [@altintas2019sleep; @exelmans2015sleep; @higuchi2005effects; @king2013impact; @peracchia2018exposure]. Late-night gaming has been shown to disrupt sleep patterns, reduce sleep duration, lower sleep quality, and increase daytime sleepiness [@exelmans2015sleep; @han2024electronic; @kristensen2021problematic]. This is especially concerning given the far-reaching effects of sleep disturbances on cognitive and emotional functioning [@cain2010electronic; @lebourgeois2017digital; @mccoy2011cognitive; @simon2020sleep; @vriend2013manipulating]. For instance, habitual gaming between 10 p.m. and 6 a.m. has been associated with an increased risk of depressive symptoms, partially mediated by daytime sleepiness [@lemola2011habitual]. Understanding the consequences of late-night gaming is thus vital for both gamers and health professionals.

## Mechanisms Linking Late-Night Gaming to Sleep Disturbance

Two key mechanisms have been proposed to explain the impact of late-night digital engagement—including gaming—on sleep. The first is the **displacement hypothesis**, which argues that late-night gaming is more harmful than daytime gaming because it cuts into sleep time [@twenge2019more; @williams2008who]. Gamers often feel compelled to continue playing and struggle with self-regulation, which can lead to insufficient sleep [@king2009understanding; @pirrone2024why; @spada2017metacognitions]. For example, adolescents experiencing a heightened sense of "flow" during challenging games delayed bedtime by up to 90 minutes [@smith2017mechanisms].

The second mechanism involves **arousal-related disturbances** in sleep architecture caused by late-night gaming. Empirical studies have shown that extended gaming, especially when involving violent content, significantly decreases REM sleep and total sleep time [@king2013impact]. @weaver2010effect highlighted that increased arousal levels due to pre-sleep gaming extend sleep latency and alter the natural progression into sleep stages. This delay in sleep onset could be exacerbated by lower melatonin levels following an evening of gaming, compared to neutral activities like board games, which are crucial for regulating the sleep-wake cycle [@hartmann2019effects].

## The Moderating Role of Chronotype

Negative effects of late-night gaming are often compounded among individuals with an eveningness chronotype—a group naturally predisposed to staying up late. Problematic gamers, who frequently possess this chronotype, are especially vulnerable to the detrimental effects of late-night gaming on sleep [@kristensen2021problematic]. Pre-sleep technology use may exacerbate the misalignment between their biological clock and societal demands by delaying sleep onset and reducing sleep duration, leading to poorer sleep quality and increased daytime sleepiness.

Research has linked evening chronotype in adolescents to greater technology use at bedtime, in turn associated with delayed sleep onset, shorter sleep duration, and poorer sleep quality [@bruni2015technology; @gumport2021impact; @kortesoja2023latenight; @reardon2023adolescent]. Additionally, while @reardon2023adolescent found that shorter sleep on weekdays was associated with greater psychological distress, technology medium and chronotype were not direct predictors of distress. @gumport2021impact found that technology use improved emotional, social, cognitive, and physical health but worsened behavioral health, measured by the consumption of junk food, caffeine, alcohol, tobacco, and other substances, in evening-type adolescents. It remains unclear how strongly these findings apply to young adults and adults, as most research has focused on adolescent populations. This leaves an open question about the extent to which evening chronotypes in older age groups are similarly affected by pre-sleep technology use.

## The Present Study

In sum, the literature indicates that video gaming, particularly when it occurs late at night, has significant implications for sleep quality, sleep duration, and overall wellbeing. This disruption can be attributed to both the displacement hypothesis [@twenge2019more; @williams2008who] and arousal-related disturbances in sleep architecture [@king2013impact]. Individual differences, such as chronotype, may moderate these effects, with eveningness chronotypes particularly vulnerable to the negative consequences of late-night gaming [@kristensen2021problematic]. The present study aims to empirically test the following hypotheses regarding the relationship between late-night gaming and sleep outcomes:

**H1**: Late-night gaming is associated with:

- **H1a**: Poorer sleep quality
- **H1b**: Shorter sleep duration
- **H1c**: Higher daytime sleepiness
- **H1d**: Lower wellbeing

In addition to testing direct relationships between late-night gaming and various sleep-related outcomes are critical to understand, we further assess the potential moderating role of chronotype, which refers to a person's natural preference for activities during certain times of the day—morningness or eveningness. Individuals with an evening chronotype tend to stay up later and may be more inclined to engage in late-night gaming, potentially exacerbating the negative impacts on sleep and wellbeing. The combination of an evening chronotype and late-night gaming may even have a compounded effect on overall wellbeing, as both factors are independently associated with poorer mental health outcomes. Given this, we propose the following:

**H2**: Chronotype moderates the relationships between late-night gaming and all outcomes in H1 (sleep quality, sleep duration, daytime sleepiness, and wellbeing), such that these negative associations are stronger for individuals with more of an eveningness chronotype.

By examining chronotype on a continuous scale as a moderating factor, this study seeks to provide a more nuanced understanding of the potential risks associated with late-night gaming and to identify individuals who may be most vulnerable to its negative effects.

# Methods

## Data Source and Measures

The analyses reported here are part of a Stage 1 Registered Report [@ballou2024psychological] and utilize data from the Open Play dataset [@ballou2025openplay], a longitudinal study that collected multi-platform video game digital trace data alongside psychological measures from adult gamers in the UK and US over a three-month period. The study combined objective behavioral telemetry from gaming platforms with repeated self-report surveys administered biweekly across six waves. Importantly, the present analyses use only a subset of the Open Play dataset, specifically data from Nintendo, Xbox, and Steam platforms, as these provide session-level (Nintendo, Xbox) or near session-level (Steam) temporal granularity necessary for hourly aggregation of playtime to operationalize late-night gaming (23:00–06:00). The following validated measures were administered via panel surveys at multiple timepoints: Wellbeing was assessed using the Short Warwick-Edinburgh Mental Well-being Scale [SWEMWBS; @tennant2007warwick], a 7-item measure of mental wellbeing covering psychological functioning and subjective well-being over the past 2 weeks, with responses on a 5-point Likert scale ranging from "None of the time" to "All of the time" (score range: 7–35). Sleep quality and duration were assessed using the Pittsburgh Sleep Quality Index [PSQI; @buysse1989pittsburgh], a 19-item questionnaire evaluating sleep quality over the past month. The measure yields seven component scores (sleep quality, sleep latency, sleep duration, sleep efficiency, sleep disturbances, use of sleep medication, and daytime dysfunction) and a global score (range: 0–21), with scores above 5 indicating poor sleep quality. Excessive daytime sleepiness was measured using the Epworth Sleepiness Scale [ESS; @johns1991new], an 8-item scale assessing the likelihood of dozing off in various situations (score range: 0–24). Higher scores indicate greater propensity for daytime sleepiness, with scores above 10 typically indicating clinically significant excessive sleepiness. Chronotype was measured at baseline (Wave 1) using the Munich Chronotype Questionnaire [MCTQ; @roenneberg2003life]. The key metric used in this study is MSF~sc~ (Mid-Sleep on Free Days corrected for sleep debt on work days), which represents an individual's natural sleep-wake preference when not constrained by social obligations. Higher MSF~sc~ values indicate a preference for eveningness (later sleep-wake times). 

## Handling Missing Data

Missingness in the longitudinal self-report outcomes (PSQI sleep duration, PSQI item 6, Epworth Sleepiness Scale, and SWEMWBS) was addressed via multiple imputation by chained equations (MICE; `mice` v3.16.0 in R), using predictive mean matching (PMM) for all continuous/ordinal targets. We imputed data to preserve statistical power under a Missing at Random assumption conditional on rich auxiliary information. The imputation model included every analysis variable, static demographics (age, BMI, SES, region, gender recoded as male/female/other), chronotype (MSF~sc~), and dynamically derived gaming exposure summaries (overall and late-night minutes averaged over the preceding 14 and 28 days, plus an indicator for weekend surveys). To guarantee that each participant contributed a full six-wave panel to the imputation model, we first expanded the self-report data to the complete pid × wave grid and inferred survey timestamps for missing waves by aligning observed dates with wave-specific medians. Those inferred dates were then used to recompute the rolling gaming exposures so that auxiliary predictors remained non-missing even when a survey wave itself had no original timestamp. We generated 20 imputed datasets with 20 iterations each—sufficient to stabilize estimates given wave-specific missingness rates up to ~55%. Diagnostic trace, density, and strip plots confirmed well-mixed chains and plausible imputations. All regressions were fit separately in each imputed dataset and combined using Rubin’s rules.

The imputed outcomes serve as the primary analytic dataset throughout the main text. Complete-case versions of every regression (using only non-imputed observations for each outcome) were re-estimated in parallel and included in the Supplement.

# Results
 
```{r}
#| label: load-libraries
#| echo: false
#| message: false
#| warning: false
#| output: false
#| cache: false

if (!require("pacman")) install.packages("pacman")
library(pacman)

p_load(
  tidyverse, lme4, marginaleffects, glmmTMB, mice, ordinal, modelsummary,
  tinytable, lubridate, data.table, mctq, glue, parameters, performance,
  withr, report, ggdist, patchwork, ggpattern
)

# Allow optional re-run of heavy preprocessing via Quarto param
rerun_preprocessing <- isTRUE(params$rerun_preprocessing)
withr::with_envvar(
  c(SKIP_PREPROCESS_ON_SOURCE = if (rerun_preprocessing) "0" else "1"),
  source("preprocess_data.R")
)

# TOGGLE HYPOTHESES EVALUATION
eval_panel <- isTRUE(params$refit_panel)
eval_diary <- isTRUE(params$refit_diary)
eval_diary_imputed <- isTRUE(params$refit_diary_imputed)
eval_appendix <- isTRUE(params$refit_appendix)

# Load helper functions
source("helpers.R")

```

```{r}
#| label: load-data
#| echo: false
#| message: false
#| warning: false

# IMPORTANT: Run preprocess_data.R first to generate processed data files
# Rscript preprocess_data.R

# Check if processed data exists, if not, run preprocessing
if (!file.exists("data/processed/selfreport.csv.gz") ||
    !file.exists("data/processed/gaming_sessions.csv.gz")) {
  message("Processed data not found. Running preprocessing...")
  preprocess_all_data()
}

# Load preprocessed data (includes 3-step filtering: outcomes + timezone + gaming)
selfreport <- read_csv("data/processed/selfreport.csv.gz", show_col_types = FALSE)
data.intake <- read_csv("data/processed/intake_clean.csv.gz", show_col_types = FALSE)

# Load stacked imputed data for Rubin's rules (20 imputations × all pid×wave rows)
imputed_long <- readRDS("data/processed/selfreport_imputed_long.rds")

# Initialize Rubin bundle containers (populated in fit-h1/h2-models chunks)
panel_h1_rubin <- list()
panel_h2_rubin <- list()

# Valid participants are those in selfreport (passed 3-step filter)
valid_pids <- unique(selfreport$pid)

# Prepare complete-case vs imputed outcome versions
OUTCOME_VARS <- c("total_hours_sleep", "psqi_comp1_quality", "epsTotal", "wemwbs", "psqi_global")

# PSQI component variables (used in DQCs and descriptives)
PSQI_COMPONENT_VARS <- c(
  "psqi_comp1_quality",
  "psqi_comp2_latency",
  "psqi_comp3_duration",
  "psqi_comp4_efficiency",
  "psqi_comp5_problems",
  "psqi_comp6_medication",
  "psqi_comp7_tired"
)

# Create versions for convenience
# Note: For per-outcome complete case analysis, use filter_by_outcome()
# with the specific outcome variable being modeled
selfreport_imputed <- filter_by_outcome(selfreport, OUTCOME_VARS, "imputed")
selfreport_completecase <- filter_by_outcome(selfreport, OUTCOME_VARS, "completecase")

# For backward compatibility with existing code
selfreport_versions <- list(
  imputed = selfreport_imputed,
  completecase = selfreport_completecase
)

selfreport <- selfreport_imputed

# Note: selfreport already contains msf_sc_numeric and msf_sc_centered from preprocessing
# msf_sc_numeric = raw chronotype for descriptives (Table 1)
# msf_sc_centered = mean-centered chronotype for regression analyses (H2 models)

# NOTE: Gaming session data (gaming_sessions.csv.gz) is loaded in the "merge gaming data" chunk
# where it is used to dynamically calculate gaming exposure windows relative to each survey date.
# The preprocessing includes:
# - 3-step filtering (outcomes + timezone + gaming)
# - UK timezone imputation
# - Late-night gaming calculation (23:00-06:00 local time)
# - Quality filtering (duration, concurrent sessions, study period)

```



```{r}
#| label: prepare gaming data
#| echo: false
#| message: false
#| warning: false

# Load session-level gaming data (needed for data quality controls)
gaming_sessions <- read_csv("data/processed/gaming_sessions.csv.gz", show_col_types = FALSE)

# NOTE: Gaming exposure windows are already computed during imputation
# and included in selfreport.csv.gz as:
# - ln_biweekly_avg_minutes_played: average late-night minutes in preceding 14 days
# - ln_monthly_avg_minutes_played: average late-night minutes in preceding 28 days
# - total_biweekly_avg_minutes_played: total average minutes in preceding 14 days
# - total_monthly_avg_minutes_played: total average minutes in preceding 28 days
#
# These were calculated using inferred dates for wave-expanded rows during the
# imputation process, ensuring consistency with the imputation model.

gaming_inputs <- lapply(selfreport_versions, build_gaming_inputs)
gamingBiweekly <- gaming_inputs$imputed$gamingBiweekly
gamingMonthly <- gaming_inputs$imputed$gamingMonthly

```

```{r}
#| label: data-quality-controls
#| echo: false
#| message: false
#| warning: false
#| code-summary: "Positive Data Quality Controls (DQC 1, 3, 4)"

# Load report package for automated results reporting
library(report)

# Load raw survey data for DQC 1
survey_biweekly <- read_csv("data/survey_biweekly.csv.gz", show_col_types = FALSE)

# ============================================
# DQC 1: Self-Reported vs. Digital Trace Playtime
# Expected: Significant positive correlation
# ============================================

# Get survey completion dates with self-reported playtime
survey_dates <- survey_biweekly |>
  filter(pid %in% valid_pids) |>
  select(pid, wave, date, self_reported_biweekly_play) |>
  mutate(
    pid = as.character(pid),
    date = as_date(date)
  ) |>
  filter(!is.na(date), !is.na(self_reported_biweekly_play))

# Calculate digital trace playtime in 14 days preceding each survey
gaming_sessions_dt <- gaming_sessions |>
  mutate(session_date = as_date(sessionStart))

dqc1_data <- survey_dates |>
  rowwise() |>
  mutate(
    window_start = date - days(14),
    window_end = date,
    digital_trace_minutes = sum(
      gaming_sessions_dt$minutes_played[
        gaming_sessions_dt$pid == pid &
        gaming_sessions_dt$session_date >= window_start &
        gaming_sessions_dt$session_date <= window_end
      ],
      na.rm = TRUE
    )
  ) |>
  ungroup() |>
  mutate(
    digital_trace_hours = digital_trace_minutes / 60,
    self_reported_hours = self_reported_biweekly_play
  ) |>
  filter(!is.na(self_reported_hours), !is.na(digital_trace_hours))

# Pearson correlation test
cor_test_dqc1 <- cor.test(
  dqc1_data$self_reported_hours, 
  dqc1_data$digital_trace_hours,
  method = "pearson"
)
dqc1_pass <- cor_test_dqc1$p.value < 0.05 & cor_test_dqc1$estimate > 0

# Generate report text for DQC 1
dqc1_report <- report(cor_test_dqc1)
dqc1_text <- paste0(
  "Self-reported playtime correlated positively with digital trace playtime (",
  sprintf("r = %.2f, 95%% CI [%.2f, %.2f], t(%d) = %.2f, p %s",
          cor_test_dqc1$estimate,
          cor_test_dqc1$conf.int[1],
          cor_test_dqc1$conf.int[2],
          cor_test_dqc1$parameter,
          cor_test_dqc1$statistic,
          ifelse(cor_test_dqc1$p.value < 0.001, "< .001", sprintf("= %.3f", cor_test_dqc1$p.value))),
  ")."
)

# Inline-friendly summary (without t statistic)
dqc1_inline <- sprintf(
  "r = %.2f, 95%% CI [%.2f, %.2f], p %s",
  cor_test_dqc1$estimate,
  cor_test_dqc1$conf.int[1],
  cor_test_dqc1$conf.int[2],
  ifelse(cor_test_dqc1$p.value < 0.001, "< .001", sprintf("= %.3f", cor_test_dqc1$p.value))
)

# ============================================
# DQC 3: Social Jetlag vs. Daytime Sleepiness
# Expected: Significant positive correlation
# ============================================

# Calculate social jetlag from MCTQ data (wave 1 only)
survey_wave1 <- survey_biweekly |>
  filter(wave == 1) |>
  select(pid, starts_with("mctq_")) |>
  mutate(pid = as.character(pid))

dqc3_mctq <- survey_wave1 |>
  mutate(
    wd = case_when(
      mctq_wd_alarm_clock_used == "Yes" ~ 5,
      mctq_wd_alarm_clock_used == "No" ~ 0,
      TRUE ~ 5
    ),
    sprep_w = hms::as_hms(mctq_wd_sleep_onset_time),
    slat_w = lubridate::dminutes(as.numeric(mctq_wd_sleep_latency_minutes)),
    se_w = hms::as_hms(mctq_wd_wake_time),
    sprep_f = hms::as_hms(mctq_fd_sleep_onset_time),
    slat_f = lubridate::dminutes(as.numeric(mctq_fd_sleep_latency_minutes)),
    se_f = hms::as_hms(mctq_fd_wake_time),
    so_w = mctq::so(sprep_w, slat_w),
    so_f = mctq::so(sprep_f, slat_f),
    sd_w = mctq::sdu(so_w, se_w),
    sd_f = mctq::sdu(so_f, se_f),
    msw = mctq::msl(so_w, sd_w),
    msf = mctq::msl(so_f, sd_f),
    sjl = mctq::sjl_rel(msw, msf),
    sjl_hours = abs(as.numeric(sjl) / 3600)
  ) |>
  select(pid, sjl_hours)

# Get ESS from selfreport data
ess_data <- selfreport |>
  filter(wave %in% c(2, 4, 6)) |>
  group_by(pid) |>
  summarise(epsTotal = mean(epsTotal, na.rm = TRUE), .groups = "drop") |>
  mutate(pid = as.character(pid))

dqc3_data <- dqc3_mctq |>
  inner_join(ess_data, by = "pid") |>
  filter(!is.na(sjl_hours), !is.na(epsTotal))

# Spearman correlation test (one-sided)
cor_test_dqc3 <- cor.test(
  dqc3_data$sjl_hours, 
  dqc3_data$epsTotal,
  method = "spearman",
  alternative = "greater"
)
dqc3_pass <- cor_test_dqc3$p.value < 0.05 & cor_test_dqc3$estimate > 0

# Generate report text for DQC 3
dqc3_text <- paste0(
  "Social jetlag was positively associated with daytime sleepiness (",
  sprintf("Spearman's ρ = %.2f, S = %.0f, p %s, one-sided",
          cor_test_dqc3$estimate,
          cor_test_dqc3$statistic,
          ifelse(cor_test_dqc3$p.value < 0.001, "< .001", sprintf("= %.3f", cor_test_dqc3$p.value))),
  ")."
)

# Inline-friendly summary
dqc3_inline <- sprintf(
  "Spearman's ρ = %.2f, p %s, one-sided",
  cor_test_dqc3$estimate,
  ifelse(cor_test_dqc3$p.value < 0.001, "< .001", sprintf("= %.3f", cor_test_dqc3$p.value))
)

# ============================================
# DQC 4: Sleep Quality vs. Wellbeing
# Expected: Significant negative correlation
# ============================================

# Get PSQI and WEMWBS from selfreport data
dqc4_person <- selfreport |>
  filter(wave %in% c(2, 4, 6)) |>
  select(pid, psqi_comp1_quality, wemwbs) |>
  filter(!is.na(psqi_comp1_quality), !is.na(wemwbs)) |>
  group_by(pid) |>
  summarise(
    psqi_sleep_quality = mean(psqi_comp1_quality, na.rm = TRUE),
    wemwbs_mean = mean(wemwbs, na.rm = TRUE),
    .groups = "drop"
  )

# Spearman correlation test (one-sided)
cor_test_dqc4 <- cor.test(
  dqc4_person$psqi_sleep_quality, 
  dqc4_person$wemwbs_mean,
  method = "spearman",
  alternative = "less"
)
dqc4_pass <- cor_test_dqc4$p.value < 0.05 & cor_test_dqc4$estimate < 0

# Generate report text for DQC 4
dqc4_text <- paste0(
  "Sleep quality was negatively associated with wellbeing (",
  sprintf("Spearman's ρ = %.2f, S = %.0f, p %s, one-sided",
          cor_test_dqc4$estimate,
          cor_test_dqc4$statistic,
          ifelse(cor_test_dqc4$p.value < 0.001, "< .001", sprintf("= %.3f", cor_test_dqc4$p.value))),
  "), confirming that higher PSQI scores (poorer sleep) are associated with lower SWEMWBS scores (poorer wellbeing)."
)

# Inline-friendly summary
dqc4_inline <- sprintf(
  "Spearman's ρ = %.2f, p %s, one-sided",
  cor_test_dqc4$estimate,
  ifelse(cor_test_dqc4$p.value < 0.001, "< .001", sprintf("= %.3f", cor_test_dqc4$p.value))
)

# Summary: All DQCs passed?
all_dqc_pass <- dqc1_pass & dqc3_pass & dqc4_pass
```

## Data Quality Controls

Prior to hypothesis testing, we conducted three positive data quality controls (DQCs) to validate expected patterns in the data. First, self-reported playtime was significantly correlated with digital trace playtime (`r dqc1_inline`), confirming convergent validity between subjective reports and objective telemetry. Second, social jetlag showed the expected positive association with daytime sleepiness (`r dqc3_inline`), replicating established findings that circadian misalignment predicts sleepiness [@Fernandes2023jetlag; @Wu2025socialjetlag]. Third, sleep quality was negatively associated with wellbeing (`r dqc4_inline`), consistent with the well-documented relationship between sleep and mental health [@Gadie2016how]. All three DQCs passed, providing confidence in the integrity of our measures before proceeding with hypothesis testing.

<!-- Also include session integrity-related DQCs for Steam, Nintendo, and Xbox ensuring no overlapping sessions on Nintendo/Xbox (within each platform), and no impossible >60 minute intervals for Steam sessions. See Supplement for positive data quality control (DQC 2) details. -->


## Sample Demographics

```{r}
#| label: sample-attrition
#| echo: false
#| message: false
#| warning: false

library(lubridate)

# Load raw panel data to reconstruct filter steps
panel_raw_attrition <- read_csv("data/raw/survey_biweekly.csv.gz", show_col_types = FALSE) |>
  mutate(date = as_datetime(date))

n_signed_up <- nrow(data.intake)
n_completed_survey <- n_distinct(panel_raw_attrition$pid)

# Step 1: Valid outcome at wave 1
step1_pids <- panel_raw_attrition |>
  filter(wave == 1) |>
  filter(
    (!is.na(wemwbs_1) | !is.na(wemwbs_2) | !is.na(wemwbs_3) |
     !is.na(wemwbs_4) | !is.na(wemwbs_5) | !is.na(wemwbs_6) | !is.na(wemwbs_7)) |
    (!is.na(psqi_bedtime_hour) | !is.na(psqi_sleep_latency_min) |
     !is.na(psqi_overall_quality) | !is.na(psqi_daytime_enthusiasm)) |
    (!is.na(eps_reading) | !is.na(eps_tv) | !is.na(eps_public) |
     !is.na(eps_passenger) | !is.na(eps_afternoon) | !is.na(eps_talking) |
     !is.na(eps_lunch) | !is.na(eps_car))
  ) |>
  distinct(pid) |>
  pull(pid)
n_step1 <- length(step1_pids)
n_excluded_step1 <- n_completed_survey - n_step1

# Step 2: Valid timezone
tz_pids <- data.intake |>
  filter(!is.na(local_timezone) | country == "UK") |>
  pull(pid) |>
  unique()
step2_pids <- intersect(step1_pids, tz_pids)
n_step2 <- length(step2_pids)
n_excluded_step2 <- n_step1 - n_step2

# Step 3: Valid gaming session (from processed gaming data, which applies quality filters)
gaming_pids <- read_csv("data/processed/gaming_sessions.csv.gz", show_col_types = FALSE) |>
  distinct(pid) |>
  pull(pid)
step3_pids <- intersect(step2_pids, gaming_pids)
n_step3 <- length(step3_pids)
n_excluded_step3 <- n_step2 - n_step3

n_analytical <- n_step3
```

Of `r format(n_signed_up, big.mark = ",")` participants who signed up for the study, `r format(n_completed_survey, big.mark = ",")` completed at least one biweekly panel survey. The analytical sample was derived from these participants by applying a three-step filter. First, participants were required to have at least one valid outcome measure (SWEMWBS, PSQI, or ESS) at wave 1, which excluded `r n_excluded_step1` participants (*n* = `r format(n_step1, big.mark = ",")`). Second, participants needed valid timezone data (either self-reported or imputed for UK participants) to accurately classify late-night gaming sessions, which excluded a further `r n_excluded_step2` participants (*n* = `r format(n_step2, big.mark = ",")`). Third, participants were required to have contributed at least one valid gaming session during the study period, which excluded `r n_excluded_step3` additional participants, yielding a final analytical sample of *N* = `r format(n_analytical, big.mark = ",")`. @tbl-demographics presents the demographic composition of both the total survey sample and the analytical sample.

```{r}
#| label: tbl-demographics
#| code-summary: "Demographics table"
#| tbl-cap: "Sample Characteristics"

selfreport_saved <- selfreport
selfreport <- selfreport_completecase

# Load required packages
library(tinytable)
library(lubridate)

# ===== STEP 1: Define samples =====
# Total sample: all participants who completed at least one panel survey
# Process raw panel data through the same pipeline as preprocessing
# Source preprocess_data.R for compute_psqi, compute_eps_total, etc.
# Use local env + save/restore wd to avoid side effects
Sys.setenv(SKIP_PREPROCESS_ON_SOURCE = "1")
local({
  saved_wd <- getwd()
  on.exit(setwd(saved_wd))
  source(file.path(here::here(), "preprocess_data.R"), local = FALSE)
})
Sys.setenv(SKIP_PREPROCESS_ON_SOURCE = "0")

panel_raw <- read_csv("data/raw/survey_biweekly.csv.gz", show_col_types = FALSE)
total_sample <- unique(panel_raw$pid)

# NOTE: 108 panel pids have no intake record — see GitHub issue #3
pids_missing_intake <- setdiff(total_sample, unique(data.intake$pid))

# Process outcomes and chronotype for ALL survey participants
panel_processed <- panel_raw |>
  compute_psqi() |>
  compute_eps_total() |>
  compute_wemwbs() |>
  compute_mctq_msf_sc() |>
  mutate(pid = as.character(pid))

# Analytical sample: participants who passed the 3-step filter
# (valid outcome at wave 1 + timezone + gaming session)
analytical_sample <- unique(selfreport$pid)

# ===== STEP 2: Prepare person-level gaming metrics =====
gaming_sessions <- read_csv("data/processed/gaming_sessions.csv.gz", show_col_types = FALSE)

compute_gaming_metrics <- function(pids) {
  gaming_sessions |>
    filter(pid %in% pids) |>
    group_by(pid) |>
    summarise(
      total_gaming_minutes = sum(minutes_played, na.rm = TRUE),
      ln_gaming_minutes = sum(latenightMinutes, na.rm = TRUE),
      total_sessions = n(),
      ln_sessions = sum(latenight == 1, na.rm = TRUE),
      .groups = "drop"
    ) |>
    mutate(
      avg_daily_total_gaming = total_gaming_minutes / 84,
      avg_daily_ln_gaming = ln_gaming_minutes / 84,
      pct_nights_ln_gaming = 100 * ln_sessions / pmax(total_sessions, 1),
      ln_one_per_week = if_else(ln_sessions >= 12, 1, 0),
      gaming_category = case_when(
        avg_daily_ln_gaming == 0 ~ "No/rare",
        avg_daily_ln_gaming <= 10 ~ "Low",
        avg_daily_ln_gaming <= 30 ~ "Moderate",
        TRUE ~ "High"
      ),
      gaming_category = factor(gaming_category,
                              levels = c("No/rare", "Low", "Moderate", "High"))
    )
}

gaming_total <- compute_gaming_metrics(total_sample)
gaming_analytical <- compute_gaming_metrics(analytical_sample)

# ===== STEP 3: Prepare person-level outcomes =====
compute_outcomes <- function(pids, data) {
  data |>
    filter(pid %in% pids, wave %in% c(2, 4, 6)) |>
    group_by(pid) |>
    summarise(
      sleep_duration = mean(total_hours_sleep, na.rm = TRUE),
      psqi_global = mean(psqi_global, na.rm = TRUE),
      psqi_comp1 = mean(psqi_comp1_quality, na.rm = TRUE),
      psqi_comp2 = mean(psqi_comp2_latency, na.rm = TRUE),
      psqi_comp3 = mean(psqi_comp3_duration, na.rm = TRUE),
      psqi_comp4 = mean(psqi_comp4_efficiency, na.rm = TRUE),
      psqi_comp5 = mean(psqi_comp5_problems, na.rm = TRUE),
      psqi_comp6 = mean(psqi_comp6_medication, na.rm = TRUE),
      psqi_comp7 = mean(psqi_comp7_tired, na.rm = TRUE),
      ess_total = mean(epsTotal, na.rm = TRUE),
      wemwbs = mean(wemwbs, na.rm = TRUE),
      .groups = "drop"
    ) |>
    mutate(
      poor_sleep = if_else(psqi_global > 5, 1, 0, missing = NA_real_)
    )
}

# Compute outcomes from processed panel data (total) and selfreport (analytical)
outcomes_total <- compute_outcomes(total_sample, panel_processed)
outcomes_analytical <- compute_outcomes(analytical_sample, selfreport)

# Chronotype from wave 1 (baseline)
compute_chronotype <- function(pids, data) {
  data |>
    filter(pid %in% pids, wave == 1) |>
    select(pid, msf_sc = msf_sc_numeric, no_alarm_free = mctq_fd_alarm_clock_used) |>
    mutate(no_alarm_free = no_alarm_free == "No")
}

chrono_total <- compute_chronotype(total_sample, panel_processed)
chrono_analytical <- compute_chronotype(analytical_sample, selfreport)

# ===== STEP 4: Join all data sources =====
assemble_master <- function(pids, gaming, outcomes, chronotype) {
  tibble(pid = pids) |>
    left_join(data.intake, by = "pid") |>
    left_join(gaming, by = "pid") |>
    left_join(outcomes, by = "pid") |>
    left_join(chronotype, by = "pid") |>
    mutate(
      gaming_category = if_else(is.na(gaming_category),
                               factor("No/rare", levels = c("No/rare", "Low", "Moderate", "High")),
                               gaming_category),
      avg_daily_total_gaming = if_else(is.na(avg_daily_total_gaming), 0, avg_daily_total_gaming),
      avg_daily_ln_gaming = if_else(is.na(avg_daily_ln_gaming), 0, avg_daily_ln_gaming),
      pct_nights_ln_gaming = if_else(is.na(pct_nights_ln_gaming), 0, pct_nights_ln_gaming),
      ln_one_per_week = if_else(is.na(ln_one_per_week), 0, ln_one_per_week)
    ) |>
    mutate(
      gender_cat = case_when(
        gender == "Man" ~ "Man",
        gender == "Woman" ~ "Woman",
        gender %in% c("Non-binary", "Genderqueer", "Genderfluid", "Agender",
                      "Demiboy", "Demiman", "He/They") ~ "Other",
        str_detect(tolower(gender), "trans") ~ "Transgender",
        is.na(gender) ~ "Not reported",
        TRUE ~ "Other"
      )
    )
}

master_data_total <- assemble_master(total_sample, gaming_total, outcomes_total, chrono_total)
master_data <- assemble_master(analytical_sample, gaming_analytical, outcomes_analytical, chrono_analytical)

# ===== STEP 5: Build table rows =====
summary_rows <- tibble(
  Characteristic = character(),
  Total = character(),
  Analytical = character()
)

# === A. SOCIODEMOGRAPHICS & ANTHROPOMETRICS ===
summary_rows <- bind_rows(summary_rows,
  tibble(Characteristic = "**A. Sociodemographics**", Total = "", Analytical = ""))

summary_rows <- bind_rows(summary_rows,
  tibble(Characteristic = "N",
         Total = as.character(nrow(master_data_total)),
         Analytical = as.character(nrow(master_data))))

summary_rows <- bind_rows(summary_rows, add_continuous("age", "Age", 1))

summary_rows <- bind_rows(summary_rows,
  tibble(Characteristic = "Gender", Total = "", Analytical = ""))
summary_rows <- bind_rows(summary_rows, add_categorical("gender_cat",
                                                        levels_order = c("Woman", "Man", "Other")))

summary_rows <- bind_rows(summary_rows,
  tibble(Characteristic = "Region", Total = "", Analytical = ""))
summary_rows <- bind_rows(summary_rows, add_categorical("region"))

summary_rows <- bind_rows(summary_rows, add_continuous("bmi", "BMI (kg/m²)", 1))
summary_rows <- bind_rows(summary_rows, add_continuous("SES_index", "SES index", 2))

# === B. CHRONOTYPE ===
summary_rows <- bind_rows(summary_rows,
  tibble(Characteristic = "**B. Chronotype**", Total = "", Analytical = ""))

no_alarm_n_tot <- sum(master_data_total$no_alarm_free, na.rm = TRUE)
no_alarm_pct_tot <- 100 * no_alarm_n_tot / sum(!is.na(master_data_total$no_alarm_free))
no_alarm_n_ana <- sum(master_data$no_alarm_free, na.rm = TRUE)
no_alarm_pct_ana <- 100 * no_alarm_n_ana / sum(!is.na(master_data$no_alarm_free))
summary_rows <- bind_rows(summary_rows,
  tibble(Characteristic = "No alarm on free days",
         Total = sprintf("%d (%.1f%%)", no_alarm_n_tot, no_alarm_pct_tot),
         Analytical = sprintf("%d (%.1f%%)", no_alarm_n_ana, no_alarm_pct_ana)))

msf_med_tot <- median(master_data_total$msf_sc, na.rm = TRUE)
msf_iqr_tot <- IQR(master_data_total$msf_sc, na.rm = TRUE)
msf_med_ana <- median(master_data$msf_sc, na.rm = TRUE)
msf_iqr_ana <- IQR(master_data$msf_sc, na.rm = TRUE)
summary_rows <- bind_rows(summary_rows,
  tibble(Characteristic = "    MCTQ-MSF~sc~ (HH:MM)^1^",
         Total = sprintf("%s (%s)", hours_to_hhmm(msf_med_tot), hours_to_hhmm(msf_iqr_tot)),
         Analytical = sprintf("%s (%s)", hours_to_hhmm(msf_med_ana), hours_to_hhmm(msf_iqr_ana))))

# === C. GAMING BEHAVIOUR ===
summary_rows <- bind_rows(summary_rows,
  tibble(Characteristic = "**C. Gaming**", Total = "", Analytical = ""))

summary_rows <- bind_rows(summary_rows,
  add_median_iqr("avg_daily_total_gaming", "Gaming (min/day)^1^", 1))
summary_rows <- bind_rows(summary_rows,
  add_median_iqr("avg_daily_ln_gaming", "LN gaming (min/day)^1^", 1))
summary_rows <- bind_rows(summary_rows,
  add_continuous("pct_nights_ln_gaming", "% nights LN gaming", 1))

# === D. SLEEP AND WELLBEING OUTCOMES ===
summary_rows <- bind_rows(summary_rows,
  tibble(Characteristic = "**D. Outcomes**", Total = "", Analytical = ""))

summary_rows <- bind_rows(summary_rows,
  add_continuous("sleep_duration", "Sleep (h)", 1))
summary_rows <- bind_rows(summary_rows,
  add_continuous("psqi_global", "PSQI global", 1))
summary_rows <- bind_rows(summary_rows,
  add_continuous("psqi_comp1", "    Sleep quality", 1))

poor_n_tot <- sum(master_data_total$poor_sleep, na.rm = TRUE)
poor_n_ana <- sum(master_data$poor_sleep, na.rm = TRUE)
summary_rows <- bind_rows(summary_rows,
  tibble(Characteristic = "    Poor sleep (PSQI>5)",
         Total = sprintf("%d (%.1f%%)", poor_n_tot,
                         100 * poor_n_tot / sum(!is.na(master_data_total$poor_sleep))),
         Analytical = sprintf("%d (%.1f%%)", poor_n_ana,
                              100 * poor_n_ana / sum(!is.na(master_data$poor_sleep)))))

summary_rows <- bind_rows(summary_rows,
  add_continuous("ess_total", "ESS", 1))

ess_n_tot <- sum(master_data_total$ess_total > 10, na.rm = TRUE)
ess_n_ana <- sum(master_data$ess_total > 10, na.rm = TRUE)
summary_rows <- bind_rows(summary_rows,
  tibble(Characteristic = "    Excessive sleepiness (ESS>10)",
         Total = sprintf("%d (%.1f%%)", ess_n_tot,
                         100 * ess_n_tot / sum(!is.na(master_data_total$ess_total))),
         Analytical = sprintf("%d (%.1f%%)", ess_n_ana,
                              100 * ess_n_ana / sum(!is.na(master_data$ess_total)))))

summary_rows <- bind_rows(summary_rows,
  add_continuous("wemwbs", "SWEMWBS", 1))

# ===== STEP 6: Display table =====
header_rows <- which(str_detect(summary_rows$Characteristic, "^\\*\\*"))
sub_item_rows <- which(str_detect(summary_rows$Characteristic, "^    ") &
                       !str_detect(summary_rows$Characteristic, "^        "))
nested_rows <- which(str_detect(summary_rows$Characteristic, "^        "))

tbl <- summary_rows |>
  tt(
    caption = "Sample Characteristics",
    notes = "Values are M (SD) unless noted. ^1^ Mdn (IQR). LN = late-night."
  ) |>
  format_tt(j = 1, markdown = TRUE) |>
  style_tt(j = 1, align = "l") |>
  style_tt(j = 2:3, align = "r") |>
  style_tt(i = header_rows, bold = TRUE) |>
  format_tt(i = header_rows, j = 1, fn = function(x) str_remove_all(x, "\\*\\*"))

if (length(sub_item_rows) > 0) {
  tbl <- tbl |>
    style_tt(i = sub_item_rows, j = 1, indent = 1) |>
    format_tt(i = sub_item_rows, j = 1, fn = function(x) str_trim(x))
}

if (length(nested_rows) > 0) {
  tbl <- tbl |>
    style_tt(i = nested_rows, j = 1, indent = 2) |>
    format_tt(i = nested_rows, j = 1, fn = function(x) str_trim(x))
}

tbl

selfreport <- selfreport_saved
```

Self-reported sleep duration in the analytical sample was `r sprintf("%.1f", mean(master_data$sleep_duration, na.rm = TRUE))` hours (SD = `r sprintf("%.1f", sd(master_data$sleep_duration, na.rm = TRUE))`), mean PSQI sleep-quality component scores were `r sprintf("%.1f", mean(master_data$psqi_comp1, na.rm = TRUE))` (SD = `r sprintf("%.1f", sd(master_data$psqi_comp1, na.rm = TRUE))`), mean daytime sleepiness was `r sprintf("%.1f", mean(master_data$ess_total, na.rm = TRUE))` on the Epworth Sleepiness Scale (SD = `r sprintf("%.1f", sd(master_data$ess_total, na.rm = TRUE))`), and mean wellbeing was `r sprintf("%.1f", mean(master_data$wemwbs, na.rm = TRUE))` on the SWEMWBS (SD = `r sprintf("%.1f", sd(master_data$wemwbs, na.rm = TRUE))`).

```{r}
#| label: fig-raincloud
#| fig-cap: "Gaming patterns and outcome distributions. (A) Hourly playtime distribution shows average daily minutes played by hour of day, with grouped bars for weekday (solid) vs weekend (striped) and late-night hours (23:00-06:00) highlighted in red. (B) Sleep Quality shows percentage of responses across ordinal categories. (C-E) Continuous variables displayed as raincloud plots with boxplots (median and IQR) and density distributions."
#| code-summary: "Combined hourly playtime, bar chart, and raincloud plots"
#| fig-width: 12
#| fig-height: 9

selfreport_saved <- selfreport
selfreport <- selfreport_completecase

# Load required packages
library(ggdist)
library(patchwork)
library(ggpattern)

# === PANEL A: Hourly Playtime Distribution ===

# Load gaming sessions data
gaming_sessions <- read_csv("data/processed/gaming_sessions.csv.gz", show_col_types = FALSE)

# Calculate participant-day counts by weekday/weekend
participant_day_counts <- gaming_sessions |>
  mutate(
    date_local = as.Date(sessionStart_local),
    isWeekend = if_else(
      weekdays(sessionStart_local) %in% c("Friday", "Saturday"),
      "Weekend",
      "Weekday"
    )
  ) |>
  distinct(pid, date_local, isWeekend) |>
  count(isWeekend, name = "n_days")

# Aggregate by hour and calculate averages
hourly_avg <- gaming_sessions |>
  mutate(
    start_hour = hour(sessionStart_local),
    isWeekend = if_else(
      weekdays(sessionStart_local) %in% c("Friday", "Saturday"),
      "Weekend",
      "Weekday"
    )
  ) |>
  group_by(start_hour, isWeekend) |>
  summarise(
    total_minutes = sum(minutes_played, na.rm = TRUE),
    n_sessions = n(),
    .groups = "drop"
  ) |>
  left_join(participant_day_counts, by = "isWeekend") |>
  mutate(
    avg_minutes_per_day = total_minutes / n_days,
    is_latenight = if_else(
      start_hour >= 23 | start_hour < 6,
      "Late-Night (23:00-06:00)",
      "Daytime (06:00-23:00)"
    )
  )

# Create hourly playtime plot
p_hourly_playtime <- ggplot(hourly_avg, aes(x = start_hour, y = avg_minutes_per_day,
                                             fill = is_latenight, pattern = isWeekend)) +
  # Grouped bars with patterns
  geom_col_pattern(
    position = position_dodge(width = 0.9),
    alpha = 0.8,
    color = "black",
    pattern_fill = "black",
    pattern_density = 0.1,
    pattern_spacing = 0.02,
    pattern_key_scale_factor = 0.5
  ) +
  # Late-night boundary markers
  geom_vline(xintercept = 23, linetype = "dashed", color = "red", linewidth = 0.7, alpha = 0.8) +
  geom_vline(xintercept = 6, linetype = "dashed", color = "red", linewidth = 0.7, alpha = 0.8) +
  annotate("text", x = 23, y = Inf, label = "23:00", vjust = 1.5, hjust = -0.1, size = 3, color = "red") +
  annotate("text", x = 6, y = Inf, label = "06:00", vjust = 1.5, hjust = 1.1, size = 3, color = "red") +
  # Color scales
  scale_fill_manual(
    values = c("Daytime (06:00-23:00)" = "steelblue", "Late-Night (23:00-06:00)" = "#E76F51"),
    name = "Time Period",
    guide = guide_legend(override.aes = list(pattern = "none"))
  ) +
  # Pattern scales for weekday/weekend
  scale_pattern_manual(
    values = c("Weekday" = "none", "Weekend" = "stripe"),
    name = "Day Type"
  ) +
  # Axes
  scale_x_continuous(
    breaks = seq(0, 23, 2),
    labels = sprintf("%02d:00", seq(0, 23, 2)),
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  # Labels and theme
  labs(
    title = "A. Gaming Playtime by Hour of Day",
    x = "Hour of Day (Local Time)",
    y = "Avg Minutes Played"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 10, color = "gray40"),
    legend.position = "bottom",
    legend.box = "horizontal"
  )

# === PANEL B: Sleep Quality (ordinal - bar chart) ===
sleep_quality_data <- selfreport |>
  filter(!is.na(psqi_comp1_quality)) |>
  mutate(
    sleep_quality = factor(psqi_comp1_quality, 
                          levels = 0:3,
                          labels = c("Very good", "Fairly good", "Fairly bad", "Very bad"))
  ) |>
  count(sleep_quality) |>
  mutate(pct = n / sum(n) * 100)

p_sleep_quality <- ggplot(sleep_quality_data, aes(x = sleep_quality, y = pct, fill = sleep_quality)) +
  geom_col(alpha = 0.8, width = 0.7) +
  geom_text(aes(label = sprintf("%.1f%%", pct)), vjust = -0.5, size = 3.5) +
  scale_fill_manual(values = c("#66c2a5", "#a6d854", "#fee08b", "#f46d43"), guide = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "B. Sleep Quality",
    subtitle = "PSQI Item 6",
    x = "",
    y = "Percentage (%)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 9),
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 10, color = "gray40")
  )

# === PANELS C-E: Continuous variables (raincloud plots) ===
continuous_data <- selfreport |>
  dplyr::select(
    pid,
    wave,
    sleep_duration = total_hours_sleep,
    daytime_sleepiness = epsTotal,
    wellbeing = wemwbs
  ) |>
  pivot_longer(
    cols = c(sleep_duration, daytime_sleepiness, wellbeing),
    names_to = "variable",
    values_to = "value"
  ) |>
  filter(!is.na(value)) |>
  mutate(
    variable = factor(variable,
                     levels = c("sleep_duration", "daytime_sleepiness", "wellbeing"),
                     labels = c("C. Sleep Duration\n(hours)",
                               "D. Daytime Sleepiness\n(ESS: 0-24)",
                               "E. Wellbeing\n(SWEMWBS: 7-35)"))
  )

p_continuous <- ggplot(continuous_data, aes(x = 1, y = value, fill = variable)) +
  # Boxplot - left layer
  geom_boxplot(
    width = 0.15,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  # Half violin (density distribution) - right layer
  stat_slab(
    adjust = 2,
    side = "right",
    scale = 0.6,
    normalize = "groups",
    alpha = 0.7,
    position = position_nudge(x = 0.2)
  ) +
  # Facet by variable with free y-scales
  facet_wrap(~ variable, scales = "free_y", nrow = 1) +
  # Styling
  scale_fill_brewer(palette = "Set2", guide = "none") +
  labs(y = "Value", x = "") +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.text = element_text(face = "bold", size = 10),
    plot.margin = margin(t = 20, r = 5, b = 5, l = 5)
  )

# Combine plots
(p_hourly_playtime) /
(p_sleep_quality + p_continuous + plot_layout(widths = c(1, 3)))

selfreport <- selfreport_saved
```

```{r}
#| label: covariate-missingness
#| echo: false
#| message: false
#| warning: false

covariate_vars <- c("age_scaled", "bmi_scaled", "SES_index_scaled", "region", "isWeekend")

covariate_data <- gaming_inputs$completecase$gamingMonthly |>
  dplyr::select(pid, all_of(covariate_vars))

total_covariate_participants <- dplyr::n_distinct(covariate_data$pid)

covariate_missing_summary <- covariate_data |>
  group_by(pid) |>
  summarise(across(all_of(covariate_vars), ~all(is.na(.x))), .groups = "drop") |>
  summarise(across(all_of(covariate_vars), ~sum(.))) |>
  pivot_longer(cols = everything(), names_to = "covariate", values_to = "n_missing") |>
  mutate(
    pct_missing = n_missing / total_covariate_participants,
    pretty_label = recode(
      covariate,
      age_scaled = "Age",
      bmi_scaled = "BMI",
      SES_index_scaled = "SES index",
      region = "Region",
      isWeekend = "Weekend indicator"
    ),
    inline = if_else(
      is.na(pct_missing),
      "n/a",
      sprintf("%d (%.1f%%)", n_missing, 100 * pct_missing)
    )
  )

covariate_missing_counts <- setNames(
  covariate_missing_summary$n_missing,
  covariate_missing_summary$covariate
)
covariate_missing_pct <- setNames(
  covariate_missing_summary$pct_missing,
  covariate_missing_summary$covariate
)
covariate_missing_inline <- setNames(
  covariate_missing_summary$inline,
  covariate_missing_summary$covariate
)
```

Missingness in the adjustment covariates was limited in the original (pre-imputation) analytical dataset: out of `r total_covariate_participants` participants, age was missing for `r covariate_missing_inline[['age_scaled']]`, BMI for `r covariate_missing_inline[['bmi_scaled']]`, SES index for `r covariate_missing_inline[['SES_index_scaled']]`, region for `r covariate_missing_inline[['region']]`, and the weekend/weekday indicator for `r covariate_missing_inline[['isWeekend']]`.


## Panel

### H1

The preregistered analyses in the Stage 1 protocol [@ballou2024psychological] specified four multilevel models in which late-night gaming minutes, averaged over 28 days (monthly) or 14 days (biweekly), predicted sleep quality (H1a), sleep duration (H1b), wellbeing (H1c), and daytime sleepiness (H1d), with random intercepts and random slopes for the late-night gaming exposure by participant and an additional random intercept for gender in the linear models. When applying this specification to the Open Play data, the preregistered random-slope structures led to convergence problems and boundary estimates (near-zero variance components), particularly for the cumulative link mixed model. To obtain stable and interpretable estimates we simplified the random-effects structure to random intercepts for participants (and for gender where supported), dropped gender as a fixed-effect covariate, and used the multiply imputed outcomes as our primary analytic dataset rather than the incomplete original outcomes. Each model was fit separately to each of the 20 imputed datasets, and parameter estimates were pooled using Rubin's rules [@rubin1987multiple], which correctly accounts for both within- and between-imputation variability.

More concretely, H1a replaces the preregistered random intercept–slope structure `(1 + monthly_avg_minutes_played | pid)` with a random intercept for participants `(1 | pid)`. For H1b, attempts to retain the preregistered random intercept for gender led to non-convergence and boundary estimates, so we used a single random intercept for participants `(1 | pid)` while keeping the same 28-day late-night average exposure. H1c and H1d follow the same logic of dropping the preregistered random slopes for participants while preserving a random intercept for gender and the same fixed-effect adjustment set (excluding gender as a fixed effect).

```{r}
#| label: h1-main-effects
#| echo: false
#| message: false
#| warning: false

# Load Rubin bundles for inline effects (either from current fit or cached)
panel_h1a_bundle <- if (length(panel_h1_rubin)) panel_h1_rubin[["H1a: Sleep Quality"]] else
  tryCatch(readRDS("output/models/imputed_h1a_rubin.rds"), error = function(e) NULL)
panel_h1b_bundle <- if (length(panel_h1_rubin)) panel_h1_rubin[["H1b: Sleep Duration"]] else
  tryCatch(readRDS("output/models/imputed_h1b_rubin.rds"), error = function(e) NULL)
panel_h1c_bundle <- if (length(panel_h1_rubin)) panel_h1_rubin[["H1c: Wellbeing"]] else
  tryCatch(readRDS("output/models/imputed_h1c_rubin.rds"), error = function(e) NULL)
panel_h1d_bundle <- if (length(panel_h1_rubin)) panel_h1_rubin[["H1d: Daytime Sleepiness"]] else
  tryCatch(readRDS("output/models/imputed_h1d_rubin.rds"), error = function(e) NULL)

h1a_ln_minutes <- get_pooled_effect(panel_h1a_bundle, "ln_monthly_avg_minutes_played_10")
h1b_ln_minutes <- get_pooled_effect(panel_h1b_bundle, "ln_monthly_avg_minutes_played_10")
h1c_ln_minutes <- get_pooled_effect(panel_h1c_bundle, "ln_biweekly_avg_minutes_played_10")
h1d_ln_minutes <- get_pooled_effect(panel_h1d_bundle, "ln_monthly_avg_minutes_played_10")

# Marginal probability computation for H1a (probit CLMM)
# Average threshold and sigma_u across 20 imputed fits, use pooled b and SE
h1a_marginal <- tryCatch({
  pv <- pool_vcov(panel_h1a_bundle$fits)
  b_gaming <- pv$qbar["ln_monthly_avg_minutes_played_10"]
  se_gaming <- sqrt(pv$vcov["ln_monthly_avg_minutes_played_10",
                             "ln_monthly_avg_minutes_played_10"])
  alpha_j <- average_re_quantity(panel_h1a_bundle$fits, function(f)
    f$alpha["Fairly good|Fairly bad"])
  sigma_u <- average_re_quantity(panel_h1a_bundle$fits, function(f)
    f$ST$pid[1, 1])
  marginal_sd <- sqrt(1 + sigma_u^2)
  pp_per_10min <- dnorm((b_gaming * 0 - alpha_j) / marginal_sd) *
    b_gaming / marginal_sd * 100
  list(pp_per_10min = pp_per_10min)
}, error = function(e) NULL)
```

Under these implemented models, the late-night gaming term in the sleep-quality model (H1a) is estimated as `r h1a_ln_minutes` (probit coefficient per 10 min/day). On the probability scale, each additional 10 minutes of average daily late-night gaming is associated with an approximate `r if (!is.null(h1a_marginal)) sprintf("%.1f", h1a_marginal$pp_per_10min)` percentage-point increase in the marginal probability of reporting fairly bad or very bad sleep quality. @fig-latenight-sleepquality-exceedance translates this coefficient to the probability scale: marginal predicted probabilities of poor sleep quality (Fairly bad or Very bad) were computed for each of the 20 imputed datasets and pooled via Rubin's rules, with 95% confidence intervals derived from the combined within- and between-imputation variance on the probability scale using the delta method. The corresponding effects are `r h1b_ln_minutes` for sleep duration (H1b), suggesting little systematic association between late-night gaming and self-reported sleep hours, `r h1c_ln_minutes` for wellbeing (H1c), indicating minimal change in SWEMWBS scores with greater late-night play, and `r h1d_ln_minutes` for daytime sleepiness (H1d), again suggesting negligible variation in Epworth scores as a function of late-night gaming. All imputed-data estimates are pooled across 20 multiply imputed datasets using Rubin's rules [@rubin1987multiple]. These models additionally adjust for age, BMI, SES index, region, and weekend versus weekday timing (with gender modelled as a random intercept in H1c and H1d), and they use rolling 14- and 28-day windows of late-night play anchored to each survey date. Full coefficient estimates, confidence intervals, and variance components are reported in the H1 regression summary table, to which we refer for all remaining parameters.



```{r}
#| label: fit-h1-models
#| code-summary: "Fit H1 models for both outcome versions"
#| cache: false
#| results: asis

h1_models <- list()
panel_h1_rubin <- list()  # Rubin's rules bundles for imputed models
if (!dir.exists("output/models")) {
  dir.create("output/models", recursive = TRUE)
}

# --- Shared prep/fit functions for H1 ---
panel_h1_prep_monthly <- function(d) {
  inputs <- build_gaming_inputs(d)
  inputs$gamingMonthly |>
    enforce_monthly_wave_subset() |>
    mutate(
      psqi_comp1_quality_rounded = round(psqi_comp1_quality),
      psqi_6_ord = factor(
        psqi_comp1_quality_rounded,
        levels = c(0, 1, 2, 3),
        labels = c("Very good", "Fairly good", "Fairly bad", "Very bad"),
        ordered = TRUE
      )
    ) |>
    filter(!is.na(psqi_6_ord))
}

panel_h1_prep_monthly_lmer <- function(d) {
  inputs <- build_gaming_inputs(d)
  inputs$gamingMonthly |> enforce_monthly_wave_subset()
}

panel_h1_prep_biweekly <- function(d) {
  inputs <- build_gaming_inputs(d)
  inputs$gamingBiweekly
}

# --- Complete-case fitting (single fit, unchanged) ---
fit_panel_h1_completecase <- function(inputs, version_label) {
  gamingMonthly <- inputs$gamingMonthly |>
    enforce_monthly_wave_subset() |>
    mutate(
      psqi_comp1_quality_rounded = round(psqi_comp1_quality),
      psqi_6_ord = factor(
        psqi_comp1_quality_rounded,
        levels = c(0, 1, 2, 3),
        labels = c("Very good", "Fairly good", "Fairly bad", "Very bad"),
        ordered = TRUE
      )
    ) |>
    filter(!is.na(psqi_6_ord))

  gamingBiweekly <- inputs$gamingBiweekly

  model.h1a <- clmm(
    psqi_6_ord ~ ln_monthly_avg_minutes_played_10 +
      age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend +
      (1 | pid),
    data = gamingMonthly,
    link = "probit",
    Hess = TRUE,
    control = clmm.control(maxIter = 5000, gradTol = 1e-3)
  )

  model.h1b <- lmer(
    total_hours_sleep ~ ln_monthly_avg_minutes_played_10 +
      age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend +
      (1 | pid),
    data = gamingMonthly,
    control = lmerControl(optimizer = "bobyqa")
  )

  model.h1c <- lmer(
    wemwbs ~ ln_biweekly_avg_minutes_played_10 +
      age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend +
      (1 | pid) + (1 | gender),
    data = gamingBiweekly,
    control = lmerControl(optimizer = "bobyqa")
  )

  model.h1d <- lmer(
    epsTotal ~ ln_monthly_avg_minutes_played_10 +
      age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend +
      (1 | pid) + (1 | gender),
    data = gamingMonthly,
    control = lmerControl(optimizer = "bobyqa")
  )

  saveRDS(model.h1a, glue("output/models/{version_label}_model_h1a.rds"))
  saveRDS(model.h1b, glue("output/models/{version_label}_model_h1b.rds"))
  saveRDS(model.h1c, glue("output/models/{version_label}_model_h1c.rds"))
  saveRDS(model.h1d, glue("output/models/{version_label}_model_h1d.rds"))

  list(
    `H1a: Sleep Quality` = model.h1a,
    `H1b: Sleep Duration` = model.h1b,
    `H1c: Wellbeing` = model.h1c,
    `H1d: Daytime Sleepiness` = model.h1d
  )
}

if (eval_panel) {
  # --- Imputed: Rubin's rules (fit on each of 20 imputations, then pool) ---
  message("Fitting H1 models on 20 imputed datasets (Rubin's rules)...")

  panel_h1_rubin[["H1a: Sleep Quality"]] <- fit_on_imputations(
    imputed_long,
    prep_fn = panel_h1_prep_monthly,
    fit_fn = function(d) clmm(
      psqi_6_ord ~ ln_monthly_avg_minutes_played_10 +
        age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend +
        (1 | pid),
      data = d, link = "probit", Hess = TRUE,
      control = clmm.control(maxIter = 5000, gradTol = 1e-3)
    )
  )

  panel_h1_rubin[["H1b: Sleep Duration"]] <- fit_on_imputations(
    imputed_long,
    prep_fn = panel_h1_prep_monthly_lmer,
    fit_fn = function(d) lmer(
      total_hours_sleep ~ ln_monthly_avg_minutes_played_10 +
        age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend +
        (1 | pid),
      data = d, control = lmerControl(optimizer = "bobyqa")
    )
  )

  panel_h1_rubin[["H1c: Wellbeing"]] <- fit_on_imputations(
    imputed_long,
    prep_fn = panel_h1_prep_biweekly,
    fit_fn = function(d) lmer(
      wemwbs ~ ln_biweekly_avg_minutes_played_10 +
        age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend +
        (1 | pid) + (1 | gender),
      data = d, control = lmerControl(optimizer = "bobyqa")
    )
  )

  panel_h1_rubin[["H1d: Daytime Sleepiness"]] <- fit_on_imputations(
    imputed_long,
    prep_fn = panel_h1_prep_monthly_lmer,
    fit_fn = function(d) lmer(
      epsTotal ~ ln_monthly_avg_minutes_played_10 +
        age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend +
        (1 | pid) + (1 | gender),
      data = d, control = lmerControl(optimizer = "bobyqa")
    )
  )

  # Save Rubin bundles for figures/inline use
  for (nm in names(panel_h1_rubin)) {
    tag <- tolower(gsub(".*: ", "h1", nm)) |> gsub(" ", "", x = _)
    tag <- switch(nm,
      "H1a: Sleep Quality" = "h1a",
      "H1b: Sleep Duration" = "h1b",
      "H1c: Wellbeing" = "h1c",
      "H1d: Daytime Sleepiness" = "h1d"
    )
    saveRDS(panel_h1_rubin[[nm]], glue("output/models/imputed_{tag}_rubin.rds"))
  }

  # For modelsummary: pooled mipo objects
  h1_models[["imputed"]] <- lapply(panel_h1_rubin, `[[`, "pooled")

} else {
  # Load cached models
  # Imputed: try Rubin bundles first
  rubin_paths <- list(
    "H1a: Sleep Quality" = "output/models/imputed_h1a_rubin.rds",
    "H1b: Sleep Duration" = "output/models/imputed_h1b_rubin.rds",
    "H1c: Wellbeing" = "output/models/imputed_h1c_rubin.rds",
    "H1d: Daytime Sleepiness" = "output/models/imputed_h1d_rubin.rds"
  )
  if (all(file.exists(unlist(rubin_paths)))) {
    panel_h1_rubin <- lapply(rubin_paths, readRDS)
    h1_models[["imputed"]] <- lapply(panel_h1_rubin, `[[`, "pooled")
  } else {
    # Fall back to old single-fit cached models
    cached <- load_h1_models_from_disk("imputed")
    if (!is.null(cached)) h1_models[["imputed"]] <- cached
  }
}

# --- Complete-case H1 (appendix) ---
if (eval_appendix) {
  h1_models[["completecase"]] <- fit_panel_h1_completecase(
    gaming_inputs[["completecase"]], "completecase"
  )
} else {
  cached_cc <- load_h1_models_from_disk("completecase")
  if (!is.null(cached_cc)) h1_models[["completecase"]] <- cached_cc
}

if (!length(h1_models)) {
  message("No cached H1 models were found. Render with `-P refit_panel:true` to create them.")
}
```


:::: {.place arguments='top, scope: "parent", float: true'}

```{r}
#| label: tbl-h1-combined
#| code-summary: "Combined summary table for all H1 models"
#| tbl-cap: "Summary of H1 Hypotheses: Effects of Late-Night Gaming on Sleep and Wellbeing"
#| output: asis


library(tinytable)

coef_map_h1 <- c(
  "ln_monthly_avg_minutes_played_10" = "Daily LN gaming (per 10 min/day, monthly)",
  "ln_biweekly_avg_minutes_played_10" = "Daily LN gaming (per 10 min/day, biweekly)",
  "age_scaled" = "Age (scaled)",
  "bmi_scaled" = "BMI (scaled)",
  "SES_index_scaled" = "SES (scaled)",
  "regionUS" = "Region: US",
  "isWeekend" = "Day: Weekend",
  "SD (Intercept pid)" = "SD (Intercept | Participant)",
  "SD (ln_monthly_avg_minutes_played pid)" = "SD (LN gaming | Participant)",
  "Cor (Intercept~ln_monthly_avg_minutes_played pid)" = "Cor (Intercept ~ LN gaming | Participant)",
  "SD (ln_biweekly_avg_minutes_played pid)" = "SD (LN gaming | Participant)",
  "Cor (Intercept~ln_biweekly_avg_minutes_played pid)" = "Cor (Intercept ~ LN gaming | Participant)",
  "SD (Observations)" = "SD (Residual)",
  "SD (Intercept gender)" = "SD (Intercept | Gender)"
)

main_version <- "imputed"
supp_versions <- setdiff(names(h1_models), main_version)


dir.create("output/tables", showWarnings = FALSE, recursive = TRUE)

# Use pooled custom rows if Rubin bundles are available, else fall back to single-fit
h1_add_rows <- if (length(panel_h1_rubin) > 0) {
  get_custom_rows_pooled(panel_h1_rubin)
} else {
  get_custom_rows(h1_models[[main_version]])
}

tt_h1 <- modelsummary(
    h1_models[[main_version]],
    fmt = fmt_sprintf("%.2f"),
    estimate  = "{estimate} [{conf.low}, {conf.high}]{stars}",
    statistic = NULL,
    gof_omit = "AIC|BIC|Log|RMSE|R2|ICC|Num|nobs",
    coef_map = coef_map_h1,
    add_rows = h1_add_rows,
    notes = c("+ p < 0.10, * p < 0.05, ** p < 0.01, *** p < 0.001",
              "LN = late-night. Confidence intervals shown in brackets.",
              "ICC = Intraclass Correlation Coefficient (adjusted)."),
    title = "H1: Effects of Late-Night Gaming on Sleep and Wellbeing (Imputed data)",
  )
tryCatch(
  save_tt(tt_h1, "output/tables/tbl-h1-combined.png", overwrite = TRUE),
  error = function(e) message("Note: PNG export skipped (", e$message, ")")
)
tt_h1


if (length(supp_versions) > 0) {
  dir.create("output/supplement", showWarnings = FALSE, recursive = TRUE)
  for (ver in supp_versions) {
    supp_path <- glue("output/supplement/h1_{ver}_summary.html")
    modelsummary(
      h1_models[[ver]],
      fmt = fmt_sprintf("%.2f"),
      estimate  = "{estimate} [{conf.low}, {conf.high}]{stars}",
      statistic = NULL,
      gof_omit = "AIC|BIC|Log|RMSE|R2|ICC|Num|nobs",
      coef_map = coef_map_h1,
      add_rows = get_custom_rows(h1_models[[ver]]),
      notes = c("+ p < 0.10, * p < 0.05, ** p < 0.01, *** p < 0.001",
                "LN = late-night. Confidence intervals shown in brackets.",
                "ICC = Intraclass Correlation Coefficient (adjusted)."),
      title = glue("H1: Effects of Late-Night Gaming on Sleep and Wellbeing ({stringr::str_to_title(ver)} data)"),
      output = supp_path
    )
  }
}
```

::::

```{r}
#| label: fig-latenight-sleepquality-exceedance
#| fig-cap: "Marginal predicted probability of poor sleep quality (Fairly bad or Very bad) as a function of late-night gaming. Predictions are derived from the H1a probit cumulative link mixed model fitted separately on each of 20 multiply imputed datasets; predicted probabilities and their within-imputation variances are then pooled via Rubin's rules. The solid line shows the pooled point estimate and the shaded ribbon the 95% confidence interval, both computed on the probability scale using the delta method applied to each imputation's threshold, gaming coefficient, and variance--covariance matrix, then combined with Rubin's within- and between-imputation variance components. Probabilities are population-average (marginalised over the participant random intercept) with other covariates held at their reference or mean values. The top panel shows the marginal density of late-night gaming in the sample; vertical reference lines mark the median, mean, 75th, and 90th percentiles."
#| fig-width: 9
#| fig-height: 6
#| code-summary: "Probability of poor sleep quality by late-night gaming"

library(ordinal)
library(patchwork)

# Load the Rubin bundle for H1a (or fall back to cached)
panel_h1a_fig_bundle <- if (!is.null(panel_h1a_bundle)) panel_h1a_bundle else
  tryCatch(readRDS("output/models/imputed_h1a_rubin.rds"), error = function(e) NULL)

fits <- panel_h1a_fig_bundle$fits
m    <- length(fits)

# Prediction grid: 0–90 min in 10-min units
ln_sequence_10 <- seq(0, 9, by = 0.1)
ln_bins <- c(0, 30, 60, 90)

# Distribution statistics for reference lines
ln_stats <- gamingMonthly |>
  filter(!is.na(ln_monthly_avg_minutes_played)) |>
  pull(ln_monthly_avg_minutes_played) |>
  {\(x) list(
    mean = mean(x),
    median = median(x),
    p75 = quantile(x, 0.75),
    p90 = quantile(x, 0.90)
  )}()

# --- Per-imputation marginal predictions with delta-method variance ----------
# For each of the 20 imputed fits, compute:
#   P(Y > "Fairly good" | x) = 1 - Phi(g(x) / denom)
# where g(x) = alpha_j - b*x, denom = sqrt(1 + sigma_u^2)
# and Var(prob) via delta method using that fit's vcov.
pred_list <- lapply(seq_len(m), function(i) {
  f <- fits[[i]]

  alpha_j  <- f$alpha["Fairly good|Fairly bad"]
  b_gaming <- coef(f)["ln_monthly_avg_minutes_played_10"]
  sigma_u  <- f$ST$pid[1, 1]
  denom    <- sqrt(1 + sigma_u^2)

  V <- as.matrix(vcov(f))
  var_alpha <- V["Fairly good|Fairly bad", "Fairly good|Fairly bad"]
  var_b     <- V["ln_monthly_avg_minutes_played_10",
                  "ln_monthly_avg_minutes_played_10"]
  cov_ab    <- V["Fairly good|Fairly bad",
                  "ln_monthly_avg_minutes_played_10"]

  g_x    <- alpha_j - b_gaming * ln_sequence_10
  prob   <- pnorm(g_x / denom, lower.tail = FALSE)

  # Within-imputation variance on probability scale (delta method)
  var_gx   <- var_alpha + ln_sequence_10^2 * var_b - 2 * ln_sequence_10 * cov_ab
  dphi     <- dnorm(g_x / denom) / denom
  var_prob <- dphi^2 * var_gx

  tibble(x = ln_sequence_10, prob = prob, var_prob = var_prob)
})

# --- Pool predictions via Rubin's rules --------------------------------------
prob_mat <- do.call(cbind, lapply(pred_list, \(d) d$prob))
var_mat  <- do.call(cbind, lapply(pred_list, \(d) d$var_prob))

qbar      <- rowMeans(prob_mat)                    # pooled point estimate
ubar      <- rowMeans(var_mat)                     # mean within-imputation var
b_var     <- apply(prob_mat, 1, var)               # between-imputation var
total_var <- ubar + (1 + 1/m) * b_var              # Rubin's total variance
se_pooled <- sqrt(total_var)

plot_data <- tibble(
  ln_monthly_avg_minutes_played = ln_sequence_10 * 10,
  exc.prob  = qbar,
  asymp.LCL = qbar - 1.96 * se_pooled,
  asymp.UCL = qbar + 1.96 * se_pooled
)

# Prepare raw data for marginal distribution
marginal_data <- gamingMonthly |>
  filter(!is.na(ln_monthly_avg_minutes_played)) |>
  select(ln_monthly_avg_minutes_played)

# Create distribution panel (top)
p_dist <- ggplot(marginal_data, aes(x = ln_monthly_avg_minutes_played)) +
  geom_density(fill = "gray70", color = "gray50", alpha = 0.6) +
  geom_vline(xintercept = ln_stats$median, linetype = "dotted",
             color = "gray40", linewidth = 0.5) +
  geom_vline(xintercept = ln_stats$mean, linetype = "dashed",
             color = "gray40", linewidth = 0.5) +
  geom_vline(xintercept = ln_stats$p75, linetype = "dotted",
             color = "gray40", linewidth = 0.5) +
  geom_vline(xintercept = ln_stats$p90, linetype = "dashed",
             color = "gray40", linewidth = 0.5) +
  scale_x_continuous(
    breaks = ln_bins,
    labels = paste0(ln_bins, " min"),
    limits = c(0, 90)
  ) +
  labs(y = "Density", x = NULL) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    plot.margin = margin(5, 5, 0, 5)
  )

# Create main prediction plot
p_main <- ggplot(plot_data, aes(x = ln_monthly_avg_minutes_played, y = exc.prob)) +
  geom_vline(xintercept = ln_stats$median, linetype = "dotted",
             color = "gray40", linewidth = 0.5) +
  geom_vline(xintercept = ln_stats$mean, linetype = "dashed",
             color = "gray40", linewidth = 0.5) +
  geom_vline(xintercept = ln_stats$p75, linetype = "dotted",
             color = "gray40", linewidth = 0.5) +
  geom_vline(xintercept = ln_stats$p90, linetype = "dashed",
             color = "gray40", linewidth = 0.5) +
  annotate("text", x = ln_stats$median, y = max(plot_data$asymp.UCL),
           label = "Median", angle = 90, vjust = -0.5, size = 3, color = "gray40") +
  annotate("text", x = ln_stats$mean, y = max(plot_data$asymp.UCL),
           label = "Mean", angle = 90, vjust = -0.5, size = 3, color = "gray40") +
  annotate("text", x = ln_stats$p75, y = max(plot_data$asymp.UCL),
           label = "75th %ile", angle = 90, vjust = -0.5, size = 3, color = "gray40") +
  annotate("text", x = ln_stats$p90, y = max(plot_data$asymp.UCL),
           label = "90th %ile", angle = 90, vjust = -0.5, size = 3, color = "gray40") +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL),
              alpha = 0.3, fill = "#d62828") +
  geom_line(linewidth = 1, color = "#d62828") +
  scale_x_continuous(
    breaks = ln_bins,
    labels = paste0(ln_bins, " min"),
    limits = c(0, 90)
  ) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = "Average Daily Late-Night Gaming (min/day)",
    y = "Predicted Probability of Poor Sleep Quality"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    plot.margin = margin(0, 5, 5, 5)
  )

# Combine plots using patchwork
(p_dist / p_main) +
  plot_layout(heights = c(1, 3)) +
  plot_annotation(
    title = "Probability of Poor Sleep Quality by Late-Night Gaming"
  )
```

### H2

The preregistration in the Stage 1 protocol [@ballou2024psychological] also specified four moderation models in which raw monthly or biweekly late-night minutes interacted with chronotype (mean sleep-on-free days corrected for sleep debt on weekdays) to predict the same four outcomes (H2a–H2d), again with random intercepts and random slopes for late-night gaming by participant and a random intercept for gender in the linear mixed models. In the analytical sample, median chronotype was `r sprintf("%.1f", stats::median(master_data$msf_sc, na.rm = TRUE))` hours with an interquartile range of `r sprintf("%.1f", stats::IQR(master_data$msf_sc, na.rm = TRUE))` hours. When transferred to the Open Play data, the preregistered moderation structure proved too demanding: the combination of highly skewed late-night exposure, strong collinearity between chronotype and late-night play, and the cumulative link mixed model for ordinal sleep quality led to non-convergence and nearly singular variance–covariance matrices, particularly for H2a.

In the final models we therefore mean-centred chronotype (`msf_sc_centered`) and simplified the random-effects structures to random intercepts for participants (and for gender where supported), while keeping the interaction terms for all four outcomes. Relative to the preregistration, H2a drops the random slope on late-night minutes for participants and uses the centred chronotype term, H2b simplifies `(1 + monthly_avg_minutes_played | pid)` to `(1 | pid)` while retaining `(1 | gender)`, H2c already specified a random-intercept-only structure for participants and mainly deviates through the use of centred chronotype, and H2d mirrors H2b in simplifying the participant random effects from `(1 + monthly_avg_minutes_played | pid)` to `(1 | pid)` alongside the same chronotype centring. The ordinal H2a model still exhibits known identifiability issues for cumulative link mixed models with interactions, so its interaction term is interpreted cautiously and our substantive conclusions about moderation rely primarily on the linear mixed-effects models (H2b–H2d). As with H1, the full pattern of main and interaction effects, along with random-effect estimates, is presented in the H2 regression summary table.

```{r}
#| label: h2-main-effects
#| echo: false
#| message: false
#| warning: false

# Load Rubin bundles for H2 inline effects
panel_h2a_bundle <- if (length(panel_h2_rubin)) panel_h2_rubin[["H2a: Sleep Quality"]] else
  tryCatch(readRDS("output/models/imputed_h2a_rubin.rds"), error = function(e) NULL)
panel_h2b_bundle <- if (length(panel_h2_rubin)) panel_h2_rubin[["H2b: Sleep Duration"]] else
  tryCatch(readRDS("output/models/imputed_h2b_rubin.rds"), error = function(e) NULL)
panel_h2c_bundle <- if (length(panel_h2_rubin)) panel_h2_rubin[["H2c: Wellbeing"]] else
  tryCatch(readRDS("output/models/imputed_h2c_rubin.rds"), error = function(e) NULL)
panel_h2d_bundle <- if (length(panel_h2_rubin)) panel_h2_rubin[["H2d: Daytime Sleepiness"]] else
  tryCatch(readRDS("output/models/imputed_h2d_rubin.rds"), error = function(e) NULL)

# Interaction effects (primary H2 tests)
h2a_ln_minutes_msf <- get_pooled_effect(panel_h2a_bundle,
  "ln_monthly_avg_minutes_played_10:msf_sc_centered")
h2b_ln_minutes_msf <- get_pooled_effect(panel_h2b_bundle,
  "ln_monthly_avg_minutes_played_10:msf_sc_centered")
h2c_ln_minutes_msf <- get_pooled_effect(panel_h2c_bundle,
  "ln_biweekly_avg_minutes_played_10:msf_sc_centered")
h2d_ln_minutes_msf <- get_pooled_effect(panel_h2d_bundle,
  "ln_monthly_avg_minutes_played_10:msf_sc_centered")

# (Exploratory) main effects for late-night gaming and chronotype
h2a_main_ln <- get_pooled_effect(panel_h2a_bundle, "ln_monthly_avg_minutes_played_10")
h2a_main_msf <- get_pooled_effect(panel_h2a_bundle, "msf_sc_centered")

h2b_main_ln <- get_pooled_effect(panel_h2b_bundle, "ln_monthly_avg_minutes_played_10")
h2b_main_msf <- get_pooled_effect(panel_h2b_bundle, "msf_sc_centered")

h2c_main_ln <- get_pooled_effect(panel_h2c_bundle, "ln_biweekly_avg_minutes_played_10")
h2c_main_msf <- get_pooled_effect(panel_h2c_bundle, "msf_sc_centered")

h2d_main_ln <- get_pooled_effect(panel_h2d_bundle, "ln_monthly_avg_minutes_played_10")
h2d_main_msf <- get_pooled_effect(panel_h2d_bundle, "msf_sc_centered")

# Marginal probability computation for H2a (probit CLMM) — pooled estimates
h2a_marginal <- tryCatch({
  pv <- pool_vcov(panel_h2a_bundle$fits)
  b_ln <- pv$qbar["ln_monthly_avg_minutes_played_10"]
  alpha_j <- average_re_quantity(panel_h2a_bundle$fits, function(f)
    f$alpha["Fairly good|Fairly bad"])
  sigma_u <- average_re_quantity(panel_h2a_bundle$fits, function(f)
    f$ST$pid[1, 1])
  marginal_sd <- sqrt(1 + sigma_u^2)
  pp_per_10min <- dnorm((b_ln * 0 - alpha_j) / marginal_sd) *
    b_ln / marginal_sd * 100
  list(pp_per_10min = pp_per_10min)
}, error = function(e) NULL)
```

Under these implemented models, the main effect of late-night gaming on sleep quality in H2a is `r h2a_main_ln` (probit coefficient per 10 min/day), corresponding to an approximate `r if (!is.null(h2a_marginal)) sprintf("%.1f", h2a_marginal$pp_per_10min)` percentage-point increase in the probability of reporting fairly bad or worse sleep per additional 10 minutes of daily late-night gaming — closely mirroring the H1a estimate. The chronotype × late-night gaming interaction is estimated as `r h2a_ln_minutes_msf`, providing little evidence that the association between late-night gaming and sleep quality differs meaningfully across the chronotype continuum. The corresponding interaction effects are `r h2b_ln_minutes_msf` for sleep duration (H2b) and `r h2d_ln_minutes_msf` for daytime sleepiness (H2d), neither of which suggests meaningful moderation by chronotype. For wellbeing (H2c), however, the interaction was `r h2c_ln_minutes_msf`, indicating that chronotype moderates the association between late-night gaming and SWEMWBS scores. Among participants with earlier chronotypes (−1 SD, i.e. morning-oriented individuals), each additional 10 minutes of daily late-night gaming was associated with a small positive change in wellbeing (~0.4 SWEMWBS points per hour). This association diminished with increasing eveningness: at the sample-mean chronotype the estimate was roughly halved (~0.2 points per hour), and at +1 SD (evening-oriented participants) the association was near zero (~0.05 points per hour). In substantive terms, the pattern suggests that late-night gaming carries a slight positive association with wellbeing for earlier chronotypes — perhaps reflecting voluntary, enjoyable leisure that does not conflict with their sleep–wake schedule — whereas for evening types, whose late-night play may more readily encroach on sleep, the association is effectively null. Despite reaching statistical significance, all of these conditional effects remain small relative to the 7–35 SWEMWBS scale. These interaction terms constitute the primary tests of H2 and are estimated conditional on the same covariate set as in H1 (age, BMI, SES index, region, and weekend versus weekday timing, with gender as a random intercept in H2b–H2d). Full model summaries, including main effects and random-effect estimates, are reported in the H2 regression table.

Although not part of our preregistered hypotheses, the main effect of late-night gaming in the sleep-quality model (H2a) was `r h2a_main_ln` and the main effect of chronotype was `r h2a_main_msf`, indicating only modest associations between each predictor and subjective sleep quality when considered marginally. Analogous main effects were `r h2b_main_ln` and `r h2b_main_msf` for sleep duration (H2b), `r h2c_main_ln` and `r h2c_main_msf` for wellbeing (H2c), and `r h2d_main_ln` and `r h2d_main_msf` for daytime sleepiness (H2d). Of these, the H2c main effect of late-night gaming on wellbeing was also statistically significant (`r h2c_main_ln`), suggesting a small positive association with SWEMWBS scores at the mean chronotype, though the magnitude is negligible in practical terms (approximately 0.2 points per hour on the 7–35 SWEMWBS scale). These exploratory estimates are reported for completeness, but should be interpreted cautiously given their non-preregistered status. The H2c interaction is visualised in @fig-h2c-interaction in the Appendix.

```{r}
#| label: fit-h2-models
#| code-summary: "Fit H2 models for both outcome versions"
#| cache: false
#| results: asis

h2_models <- list()
panel_h2_rubin <- list()  # Rubin's rules bundles for imputed H2 models

# --- Shared prep functions for H2 ---
panel_h2_prep_monthly_clmm <- function(d) {
  inputs <- build_gaming_inputs(d)
  inputs$gamingMonthly |>
    enforce_monthly_wave_subset() |>
    mutate(
      psqi_comp1_quality_rounded = round(psqi_comp1_quality),
      psqi_6_ord = factor(
        psqi_comp1_quality_rounded,
        levels = c(0, 1, 2, 3),
        labels = c("Very good", "Fairly good", "Fairly bad", "Very bad"),
        ordered = TRUE
      )
    ) |>
    filter(!is.na(psqi_6_ord), !is.na(msf_sc_centered))
}

panel_h2_prep_monthly_lmer <- function(d) {
  inputs <- build_gaming_inputs(d)
  inputs$gamingMonthly |>
    enforce_monthly_wave_subset() |>
    filter(!is.na(msf_sc_centered), !is.na(total_hours_sleep), !is.na(epsTotal))
}

panel_h2_prep_biweekly <- function(d) {
  inputs <- build_gaming_inputs(d)
  inputs$gamingBiweekly |>
    filter(!is.na(msf_sc_centered), !is.na(wemwbs))
}

# --- Complete-case fitting (single fit, unchanged) ---
fit_panel_h2_completecase <- function(inputs, version_label) {
  gamingMonthly <- inputs$gamingMonthly |> enforce_monthly_wave_subset()
  gamingBiweekly <- inputs$gamingBiweekly

  gamingMonthly_complete <- gamingMonthly |>
    filter(!is.na(msf_sc_centered), !is.na(total_hours_sleep), !is.na(epsTotal))
  gamingBiweekly_complete <- gamingBiweekly |>
    filter(!is.na(msf_sc_centered), !is.na(wemwbs))

  if (nrow(gamingMonthly_complete) < 50 || nrow(gamingBiweekly_complete) < 50) {
    message(sprintf("Skipping H2 completecase: insufficient cases (monthly: %d, biweekly: %d)",
                    nrow(gamingMonthly_complete), nrow(gamingBiweekly_complete)))
    return(NULL)
  }

  gamingMonthly_h2 <- gamingMonthly |>
    mutate(
      psqi_comp1_quality_rounded = round(psqi_comp1_quality),
      psqi_6_ord = factor(
        psqi_comp1_quality_rounded,
        levels = c(0, 1, 2, 3),
        labels = c("Very good", "Fairly good", "Fairly bad", "Very bad"),
        ordered = TRUE
      )
    ) |>
    filter(!is.na(psqi_6_ord), !is.na(msf_sc_centered))

  model.h2a <- clmm(
    psqi_6_ord ~ ln_monthly_avg_minutes_played_10 * msf_sc_centered +
      age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend +
      (1 | pid),
    data = gamingMonthly_h2,
    link = "probit", Hess = TRUE,
    control = clmm.control(maxIter = 5000, gradTol = 1e-3)
  )

  model.h2b <- lmer(
    total_hours_sleep ~ ln_monthly_avg_minutes_played_10 * msf_sc_centered +
      age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend +
      (1 | pid) + (1 | gender),
    data = gamingMonthly_complete
  )

  model.h2c <- lmer(
    wemwbs ~ ln_biweekly_avg_minutes_played_10 * msf_sc_centered +
      age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend +
      (1 | pid) + (1 | gender),
    data = gamingBiweekly_complete
  )

  model.h2d <- lmer(
    epsTotal ~ ln_monthly_avg_minutes_played_10 * msf_sc_centered +
      age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend +
      (1 | pid) + (1 | gender),
    data = gamingMonthly_complete
  )

  saveRDS(model.h2a, glue("output/models/{version_label}_model_h2a.rds"))
  saveRDS(model.h2b, glue("output/models/{version_label}_model_h2b.rds"))
  saveRDS(model.h2c, glue("output/models/{version_label}_model_h2c.rds"))
  saveRDS(model.h2d, glue("output/models/{version_label}_model_h2d.rds"))

  list(
    `H2a: Sleep Quality` = model.h2a,
    `H2b: Sleep Duration` = model.h2b,
    `H2c: Wellbeing` = model.h2c,
    `H2d: Daytime Sleepiness` = model.h2d
  )
}

if (eval_panel) {
  # --- Imputed: Rubin's rules ---
  message("Fitting H2 models on 20 imputed datasets (Rubin's rules)...")

  panel_h2_rubin[["H2a: Sleep Quality"]] <- fit_on_imputations(
    imputed_long,
    prep_fn = panel_h2_prep_monthly_clmm,
    fit_fn = function(d) clmm(
      psqi_6_ord ~ ln_monthly_avg_minutes_played_10 * msf_sc_centered +
        age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend +
        (1 | pid),
      data = d, link = "probit", Hess = TRUE,
      control = clmm.control(maxIter = 5000, gradTol = 1e-3)
    )
  )

  panel_h2_rubin[["H2b: Sleep Duration"]] <- fit_on_imputations(
    imputed_long,
    prep_fn = panel_h2_prep_monthly_lmer,
    fit_fn = function(d) lmer(
      total_hours_sleep ~ ln_monthly_avg_minutes_played_10 * msf_sc_centered +
        age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend +
        (1 | pid) + (1 | gender),
      data = d
    )
  )

  panel_h2_rubin[["H2c: Wellbeing"]] <- fit_on_imputations(
    imputed_long,
    prep_fn = panel_h2_prep_biweekly,
    fit_fn = function(d) lmer(
      wemwbs ~ ln_biweekly_avg_minutes_played_10 * msf_sc_centered +
        age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend +
        (1 | pid) + (1 | gender),
      data = d
    )
  )

  panel_h2_rubin[["H2d: Daytime Sleepiness"]] <- fit_on_imputations(
    imputed_long,
    prep_fn = panel_h2_prep_monthly_lmer,
    fit_fn = function(d) lmer(
      epsTotal ~ ln_monthly_avg_minutes_played_10 * msf_sc_centered +
        age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend +
        (1 | pid) + (1 | gender),
      data = d
    )
  )

  for (nm in names(panel_h2_rubin)) {
    tag <- switch(nm,
      "H2a: Sleep Quality" = "h2a",
      "H2b: Sleep Duration" = "h2b",
      "H2c: Wellbeing" = "h2c",
      "H2d: Daytime Sleepiness" = "h2d"
    )
    saveRDS(panel_h2_rubin[[nm]], glue("output/models/imputed_{tag}_rubin.rds"))
  }

  h2_models[["imputed"]] <- lapply(panel_h2_rubin, `[[`, "pooled")

} else {
  # Load cached models
  rubin_paths_h2 <- list(
    "H2a: Sleep Quality" = "output/models/imputed_h2a_rubin.rds",
    "H2b: Sleep Duration" = "output/models/imputed_h2b_rubin.rds",
    "H2c: Wellbeing" = "output/models/imputed_h2c_rubin.rds",
    "H2d: Daytime Sleepiness" = "output/models/imputed_h2d_rubin.rds"
  )
  if (all(file.exists(unlist(rubin_paths_h2)))) {
    panel_h2_rubin <- lapply(rubin_paths_h2, readRDS)
    h2_models[["imputed"]] <- lapply(panel_h2_rubin, `[[`, "pooled")
  } else {
    cached <- load_h2_models_from_disk("imputed")
    if (!is.null(cached)) h2_models[["imputed"]] <- cached
  }
}

# --- Complete-case H2 (appendix) ---
if (eval_appendix) {
  cc_fitted <- fit_panel_h2_completecase(gaming_inputs[["completecase"]], "completecase")
  if (!is.null(cc_fitted)) h2_models[["completecase"]] <- cc_fitted
} else {
  cached_cc <- load_h2_models_from_disk("completecase")
  if (!is.null(cached_cc)) h2_models[["completecase"]] <- cached_cc
}

```



:::: {.place arguments='top, scope: "parent", float: true'}

```{r}
#| label: tbl-h2-combined
#| code-summary: "Combined summary table for all H2 models"
#| tbl-cap: "Summary of H2 Hypotheses: Chronotype Moderation of Late-Night Gaming Effects"
#| output: asis


  library(tinytable)

  coef_map_h2 <- c(
    "ln_monthly_avg_minutes_played_10" = "Daily LN gaming (per 10 min/day, monthly)",
    "ln_biweekly_avg_minutes_played_10" = "Daily LN gaming (per 10 min/day, biweekly)",
    "msf_sc_centered" = "Chronotype (h, centered)",
    "ln_monthly_avg_minutes_played_10:msf_sc_centered" = "LN gaming × Chronotype (h, monthly)",
    "ln_biweekly_avg_minutes_played_10:msf_sc_centered" = "LN gaming × Chronotype (h, biweekly)",
    "age_scaled" = "Age (scaled)",
    "bmi_scaled" = "BMI (scaled)",
    "SES_index_scaled" = "SES (scaled)",
    "regionUS" = "Region: US",
    "isWeekend" = "Day: Weekend",
    "SD (Intercept pid)" = "SD (Intercept | Participant)",
    "SD (Intercept gender)" = "SD (Intercept | Gender)",
    "SD (Observations)" = "SD (Residual)"
  )

  main_version <- "imputed"
  supp_versions <- setdiff(names(h2_models), main_version)

h2_add_rows <- if (length(panel_h2_rubin) > 0) {
  get_custom_rows_pooled(panel_h2_rubin)
} else {
  get_custom_rows(h2_models[[main_version]])
}

tt_h2 <- modelsummary(
  h2_models[[main_version]],
  fmt = fmt_sprintf("%.2f"),
  estimate  = "{estimate} [{conf.low}, {conf.high}]{stars}",
  statistic = NULL,
  gof_omit = "AIC|BIC|Log|RMSE|R2|ICC|Num|nobs",
  coef_map = coef_map_h2,
  add_rows = h2_add_rows,
  notes = c("+ p < 0.10, * p < 0.05, ** p < 0.01, *** p < 0.001",
            "LN = late-night. Confidence intervals shown in brackets.",
            "ICC = Intraclass Correlation Coefficient (adjusted)."),
  title = "H2: Chronotype Moderation of Late-Night Gaming Effects (Imputed data)"
)
tryCatch(
  save_tt(tt_h2, "output/tables/tbl-h2-combined.png", overwrite = TRUE),
  error = function(e) message("Note: PNG export skipped (", e$message, ")")
)
tt_h2


  if (length(supp_versions) > 0) {
    dir.create("output/supplement", showWarnings = FALSE, recursive = TRUE)
    for (ver in supp_versions) {
      if (!is.null(h2_models[[ver]])) {
        supp_path <- glue("output/supplement/h2_{ver}_summary.html")
        modelsummary(
          h2_models[[ver]],
          fmt = fmt_sprintf("%.2f"),
          estimate  = "{estimate} [{conf.low}, {conf.high}]{stars}",
          statistic = NULL,
          gof_omit = "AIC|BIC|Log|RMSE|R2|ICC|Num|nobs",
          coef_map = coef_map_h2,
          add_rows = get_custom_rows(h2_models[[ver]]),
          notes = c("+ p < 0.10, * p < 0.05, ** p < 0.01, *** p < 0.001",
                    "LN = late-night. Confidence intervals shown in brackets.",
                    "ICC = Intraclass Correlation Coefficient (adjusted)."),
          title = glue("H2: Chronotype Moderation of Late-Night Gaming Effects ({stringr::str_to_title(ver)} data)"),
          output = supp_path
        )
      }
    }
  }

```

::::

## Diary

### H1 and H2

```{r}
#| label: diary-data-prep
#| cache: true
#| echo: false
#| message: false
#| warning: false

# Load raw diary data
diary_survey_raw <- read_csv(
  "data/raw/survey_daily.csv.gz",
  show_col_types = FALSE
)

diary_timeuse_raw <- read_csv(
  "data/raw/timeuse.csv.gz",
  show_col_types = FALSE
)

diary_gaming_sessions_raw <- read_csv(
  "data/processed/gaming_sessions.csv.gz",
  show_col_types = FALSE
)

# --- Enforce ≥10h gap between consecutive diary entries per participant ---
diary_survey <- diary_survey_raw |>
  mutate(
    diary_ts = ymd_hms(date, tz = "UTC"),
    diary_day = as_date(diary_ts),
    wave_id = as.integer(wave),
    sleep_quality_num = recode_sleep_quality(sleep_diary_quality)
  ) |>
  arrange(pid, diary_ts) |>
  group_by(pid) |>
  mutate(
    gap_hours = as.numeric(difftime(diary_ts, lag(diary_ts), units = "hours")),
    keep_entry = is.na(gap_hours) | gap_hours >= 10
  ) |>
  filter(keep_entry) |>
  ungroup()

# --- Derive sleep metrics from time-use diaries ---
diary_timeuse <- diary_timeuse_raw |>
  filter(quality_flag == "Good quality") |>
  mutate(
    diary_day = ymd(date),
    wave_id = as.integer(daily_wave),
    start_ts = ymd_hms(start_time, tz = "UTC"),
    end_ts = ymd_hms(end_time, tz = "UTC")
  ) |>
  arrange(pid, wave_id, start_ts) |>
  group_by(pid, wave_id) |>
  mutate(activity_run = cumsum(activity != lag(activity, default = first(activity)))) |>
  group_by(pid, wave_id, activity_run, activity) |>
  summarise(
    diary_day = min(diary_day, na.rm = TRUE),
    start_ts = min(start_ts, na.rm = TRUE),
    end_ts = max(end_ts, na.rm = TRUE),
    .groups = "drop_last"
  ) |>
  ungroup() |>
  mutate(
    start_minutes = hour(start_ts) * 60 + minute(start_ts) + second(start_ts) / 60,
    end_minutes = hour(end_ts) * 60 + minute(end_ts) + second(end_ts) / 60,
    duration_minutes_raw = end_minutes - start_minutes,
    duration_minutes = case_when(
      is.na(start_minutes) | is.na(end_minutes) ~ NA_real_,
      duration_minutes_raw < 0 ~ duration_minutes_raw + 24 * 60,
      TRUE ~ duration_minutes_raw
    ),
    duration_hours = duration_minutes / 60
  )

diary_sleep_summary <- diary_timeuse |>
  filter(activity == "Sleep") |>
  arrange(pid, wave_id, start_ts) |>
  group_by(pid, wave_id) |>
  summarise(
    n_sleep_segments = n(),
    sleep_duration_hours = sum(duration_hours, na.rm = TRUE),
    bedtime_utc = max(start_ts, na.rm = TRUE),
    wakeup_utc = first(end_ts),
    diary_day = min(diary_day, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    n_sleep_segments = if_else(n_sleep_segments > 10, NA_integer_, n_sleep_segments),
    sleep_duration_hours = if_else(n_sleep_segments > 10, NA_real_, sleep_duration_hours),
    bedtime_utc = if_else(n_sleep_segments > 10, as.POSIXct(NA), bedtime_utc),
    wakeup_utc = if_else(n_sleep_segments > 10, as.POSIXct(NA), wakeup_utc)
  )

# Remove sleep duration outliers beyond 3 SD
diary_sleep_stats <- diary_sleep_summary |>
  summarise(mu = mean(sleep_duration_hours, na.rm = TRUE),
            sigma = sd(sleep_duration_hours, na.rm = TRUE))
if (!is.na(diary_sleep_stats$sigma) && diary_sleep_stats$sigma > 0) {
  diary_sleep_summary <- diary_sleep_summary |>
    mutate(sleep_duration_hours = if_else(
      abs(sleep_duration_hours - diary_sleep_stats$mu) > 3 * diary_sleep_stats$sigma,
      NA_real_, sleep_duration_hours
    ))
}

# --- Late-night gaming in prior 24h ---
diary_gaming_sessions <- diary_gaming_sessions_raw |>
  mutate(
    session_start = ymd_hms(sessionStart, tz = "UTC", quiet = TRUE),
    session_end = ymd_hms(sessionEnd, tz = "UTC", quiet = TRUE),
    latenightMinutes = replace_na(latenightMinutes, 0)
  ) |>
  filter(!is.na(session_start), !is.na(session_end), session_end > session_start)

diary_late_night <- diary_survey |>
  select(pid, wave_id, diary_ts) |>
  left_join(diary_gaming_sessions, by = "pid") |>
  filter(
    session_end > diary_ts - hours(24),
    session_start <= diary_ts
  ) |>
  mutate(
    overlap_minutes = pmax(0, as.numeric(difftime(
      pmin(session_end, diary_ts),
      pmax(session_start, diary_ts - hours(24)),
      units = "mins"
    ))),
    session_minutes = as.numeric(difftime(session_end, session_start, units = "mins")),
    overlap_frac = if_else(session_minutes > 0, overlap_minutes / session_minutes, 0),
    late_night_minutes_overlap = latenightMinutes * overlap_frac
  ) |>
  group_by(pid, wave_id) |>
  summarise(
    late_night_gaming_hours_24h = sum(late_night_minutes_overlap, na.rm = TRUE) / 60,
    .groups = "drop"
  )

# --- Chronotype lookup from wave 1 self-report (reuse already-loaded data) ---
diary_chronotype_lookup <- selfreport_imputed |>
  filter(wave == 1) |>
  distinct(pid, msf_sc_numeric, msf_sc_centered, mctq_fd_alarm_clock_used) |>
  mutate(
    chronotype_hours = msf_sc_numeric,
    chronotype_centered = msf_sc_centered,
    no_alarm_free = mctq_fd_alarm_clock_used == "No"
  ) |>
  select(pid, chronotype_hours, chronotype_centered, no_alarm_free)

# --- Covariate lookup from intake data (reuse already-loaded data) ---
diary_covariate_lookup <- data.intake |>
  group_by(pid) |>
  slice_head(n = 1) |>
  ungroup() |>
  transmute(
    pid,
    region = factor(country),
    gender = factor(na_if(gender, "NA")),
    age = suppressWarnings(as.numeric(age)),
    height_raw = suppressWarnings(as.numeric(height)),
    weight_raw = suppressWarnings(as.numeric(weight)),
    edu_level = na_if(edu_level, "NA"),
    employment = na_if(employment, "NA")
  ) |>
  mutate(
    height_m = case_when(
      is.na(height_raw) ~ NA_real_,
      height_raw < 100 ~ height_raw * 0.0254,
      TRUE ~ height_raw / 100
    ),
    weight_kg = case_when(
      is.na(weight_raw) ~ NA_real_,
      weight_raw > 150 ~ weight_raw * 0.45359237,
      TRUE ~ weight_raw
    ),
    bmi = weight_kg / (height_m^2),
    bmi = if_else(is.finite(bmi) & bmi >= 10 & bmi <= 70, bmi, NA_real_),
    edu_score = case_when(
      is.na(edu_level) ~ NA_real_,
      str_detect(str_to_lower(edu_level), "no formal") ~ 1,
      str_detect(str_to_lower(edu_level), "secondary|high school|gcse") ~ 2,
      str_detect(str_to_lower(edu_level), "some college|associate") ~ 3,
      str_detect(str_to_lower(edu_level), "bachelor") ~ 4,
      str_detect(str_to_lower(edu_level), "post-graduate|postgraduate|master|phd|doctor") ~ 5,
      TRUE ~ NA_real_
    ),
    emp_score = case_when(
      is.na(employment) ~ NA_real_,
      str_detect(str_to_lower(employment), "full-time|full time|self-employed|self employ|freelan|streamer|gig|military") ~ 4,
      str_detect(str_to_lower(employment), "part-time|part time") ~ 3,
      str_detect(str_to_lower(employment), "student") ~ 2,
      str_detect(str_to_lower(employment), "retired") ~ 2,
      str_detect(str_to_lower(employment), "homemaker|stay-at-home|disabled|unable to work|caregiver|not currently employed|unemployed|prolific") ~ 1,
      TRUE ~ NA_real_
    ),
    SES_index = edu_score + emp_score
  ) |>
  select(pid, age, bmi, SES_index, region, gender)

# Restrict to diary participants and scale age/bmi within this subsample
# (scaling within diary sample avoids compressed covariate ranges that cause
# numerical instability in CLMM when the diary subset differs from full intake)
diary_all_pids <- unique(diary_survey$pid)
diary_covariate_lookup <- diary_covariate_lookup |>
  filter(pid %in% diary_all_pids) |>
  mutate(
    age_scaled = as.numeric(scale(age)),
    bmi_scaled = as.numeric(scale(bmi))
  ) |>
  select(pid, age_scaled, bmi_scaled, SES_index, region, gender)

# Impute missing SES_index for diary participants using MICE (PMM, m = 5)
if (sum(is.na(diary_covariate_lookup$SES_index)) > 0) {
  set.seed(42)
  diary_ses_imp <- mice::mice(
    diary_covariate_lookup |> select(SES_index, age_scaled, bmi_scaled, gender, region),
    m = 5, method = "pmm", maxit = 10, seed = 42, printFlag = FALSE
  )
  diary_ses_complete <- mice::complete(diary_ses_imp, 1) |>
    mutate(pid = diary_covariate_lookup$pid)
  diary_covariate_lookup <- diary_covariate_lookup |>
    select(-SES_index) |>
    left_join(diary_ses_complete |> select(pid, SES_index), by = "pid")
}

diary_covariate_lookup <- diary_covariate_lookup |>
  mutate(SES_index_scaled = as.numeric(scale(SES_index))) |>
  select(pid, age_scaled, bmi_scaled, SES_index_scaled, region, gender)

# --- Assemble diary_long ---
diary_long <- diary_survey |>
  select(
    pid, wave_id,
    day = diary_day,
    playing_last_24h = played24hr,
    sleep_quality = sleep_diary_quality,
    sleep_quality_num
  ) |>
  left_join(diary_sleep_summary, by = c("pid", "wave_id")) |>
  left_join(diary_chronotype_lookup, by = "pid") |>
  left_join(diary_covariate_lookup, by = "pid") |>
  left_join(diary_late_night, by = c("pid", "wave_id")) |>
  mutate(
    late_night_gaming_hours_24h = replace_na(late_night_gaming_hours_24h, 0),
    # Friday/Saturday = weekend (wday: 1=Sun...7=Sat)
    is_weekend = lubridate::wday(day) %in% c(6, 7)
  )

# --- Build diary_ocat with within/between decomposition ---
diary_gender_levels <- sort(unique(na.omit(diary_long$gender)))
if (length(diary_gender_levels) < 2) diary_gender_levels <- c(diary_gender_levels, "other_gender")

diary_ocat <- diary_long |>
  filter(!is.na(sleep_quality_num)) |>
  mutate(
    sleep_quality_num = pmax(pmin(as.integer(sleep_quality_num), 5L), 1L),
    late_night_minutes_trim = pmin(
      late_night_gaming_hours_24h * 60,
      quantile(late_night_gaming_hours_24h * 60, 0.99, na.rm = TRUE)
    ),
    sleep_quality_ord = factor(
      sleep_quality_num,
      levels = 5:1,
      labels = c("Very good", "Good", "Fair", "Poor", "Very poor"),
      ordered = TRUE
    ),
    is_weekend = factor(as.logical(is_weekend), levels = c(FALSE, TRUE)),
    gender = factor(gender, levels = diary_gender_levels)
  ) |>
  group_by(pid) |>
  mutate(
    late_night_pid_mean = mean(late_night_minutes_trim, na.rm = TRUE),
    late_night_within = late_night_minutes_trim - late_night_pid_mean
  ) |>
  ungroup() |>
  mutate(
    late_night_between = late_night_pid_mean - mean(late_night_pid_mean, na.rm = TRUE),
    # Per-10-min scaling: effect per 10 additional minutes of late-night gaming
    ln_within_10  = late_night_within / 10,
    ln_between_10 = late_night_between / 10
  )

diary_ocat_h2 <- diary_ocat |> filter(!is.na(chronotype_centered))

diary_n_participants <- length(unique(diary_ocat$pid))
diary_n_entries <- nrow(diary_ocat)
```

```{r}
#| label: fit-diary-h1-model
#| echo: false
#| message: false
#| warning: false
#| cache: false
#| eval: !expr eval_diary

library(ordinal)

# H1: Direct effects (CLMM with random intercept for participant)
# Gender is excluded: (1|gender) with only 3 levels does not converge
# Region is excluded: all diary participants are US-only (no UK participants)
# Probit link used for convergence. Late-night gaming expressed per 10 minutes.
# Positive coefficients = higher probability of worse sleep (levels=5:1 coding).
diary_h1_formula <- sleep_quality_ord ~
  ln_within_10 + ln_between_10 +
  age_scaled + bmi_scaled + SES_index_scaled + is_weekend + (1 | pid)

diary_model_h1 <- clmm(diary_h1_formula, data = diary_ocat, link = "probit", Hess = TRUE,
                        control = clmm.control(maxIter = 50000, gradTol = 0.5))
saveRDS(diary_model_h1, "output/models/diary_model_h1_sleepquality.rds")

# Spline comparison (df = 2:6) — use CLM (no RE) for speed; linearity check only
diary_covariate_terms <- c("ln_between_10", "age_scaled", "bmi_scaled",
                           "SES_index_scaled", "region", "is_weekend")

diary_spline_fits <- purrr::map(2:6, \(df) {
  f <- as.formula(paste(
    "sleep_quality_ord ~",
    paste(c(sprintf("splines::ns(ln_within_10, df = %d)", df),
            diary_covariate_terms), collapse = " + ")
  ))
  mod <- clm(f, data = diary_ocat, link = "probit", Hess = TRUE)
  list(df = df, model = mod)
})

diary_clm_linear <- clm(
  sleep_quality_ord ~ ln_within_10 + ln_between_10 +
    age_scaled + bmi_scaled + SES_index_scaled + region + is_weekend,
  data = diary_ocat, link = "probit", Hess = TRUE
)

diary_spline_compare <- tibble(
  df = purrr::map_int(diary_spline_fits, "df"),
  AIC = purrr::map_dbl(diary_spline_fits, \(x) AIC(x$model))
) |> arrange(AIC) |> mutate(deltaAIC = AIC - min(AIC))

diary_linear_aic <- AIC(diary_clm_linear)
diary_best_spline_aic <- diary_spline_compare$AIC[1]

diary_spline_compare$linear_AIC <- diary_linear_aic
saveRDS(diary_spline_compare, "output/models/diary_spline_compare.rds")
```

```{r}
#| label: fit-diary-h2-model
#| echo: false
#| message: false
#| warning: false
#| cache: false
#| eval: !expr eval_diary

library(ordinal)

# H2: Chronotype moderation — interaction on BOTH within and between
# Probit link for convergence. Per-10min gaming × centered chronotype interaction.
# Positive coefficients = higher probability of worse sleep (levels=5:1 coding).

# Region cannot be included in H2: all participants with chronotype data (n=509)
# are from the US, so region has only one level in diary_ocat_h2
diary_h2_formula <- sleep_quality_ord ~
  ln_within_10 * chronotype_centered +
  ln_between_10 * chronotype_centered +
  age_scaled + bmi_scaled + SES_index_scaled + is_weekend + (1 | pid)

diary_model_h2 <- clmm(
  diary_h2_formula,
  data    = diary_ocat_h2,
  link    = "probit",
  Hess    = TRUE,
  control = clmm.control(maxIter = 50000, gradTol = 0.5)
)
saveRDS(diary_model_h2, "output/models/diary_model_h2_sleepquality.rds")
```

```{r}
#| label: fit-diary-imputed-models
#| echo: false
#| message: false
#| warning: false
#| cache: false
#| eval: !expr eval_diary_imputed

library(ordinal)

# Load imputed diary data (from imputation_diary.R)
diary_imputed_long_path <- "data/processed/diary_imputed_long.rds"
stopifnot(file.exists(diary_imputed_long_path))
diary_imputed_raw <- readRDS(diary_imputed_long_path)

# The imputed data has day_num (1-30) but diary models need late-night gaming
# exposure from telemetry. We join the observed late-night gaming and rebuild
# diary_ocat for each imputation.

# Prep function: builds diary_ocat-equivalent from one imputation slice
diary_imputed_prep_h1 <- function(d) {
  # Restrict to originally observed diary days only. Rows where
  # day_type_weekend_imputed_flag == TRUE are "phantom" days added during
  # grid expansion; their gaming exposure is 0 and dilutes between-person variance.
  d <- d |> filter(!day_type_weekend_imputed_flag)

  # Join late-night gaming exposure via pid + day_num → wave_id mapping
  d <- d |>
    left_join(
      diary_late_night |> rename(day_num = wave_id),
      by = c("pid", "day_num")
    ) |>
    mutate(
      late_night_gaming_hours_24h = replace_na(late_night_gaming_hours_24h, 0),
      sleep_quality_num = pmax(pmin(as.integer(sleep_diary_quality_num), 5L), 1L),
      late_night_minutes_trim = pmin(
        late_night_gaming_hours_24h * 60,
        quantile(late_night_gaming_hours_24h * 60, 0.99, na.rm = TRUE)
      ),
      sleep_quality_ord = factor(
        sleep_quality_num,
        levels = 5:1,
        labels = c("Very good", "Good", "Fair", "Poor", "Very poor"),
        ordered = TRUE
      ),
      is_weekend = factor(
        day_type_weekend == 1 | day_type_dayoff == 1,
        levels = c(FALSE, TRUE)
      )
    ) |>
    group_by(pid) |>
    mutate(
      late_night_pid_mean = mean(late_night_minutes_trim, na.rm = TRUE),
      late_night_within = late_night_minutes_trim - late_night_pid_mean
    ) |>
    ungroup() |>
    mutate(
      late_night_between = late_night_pid_mean - mean(late_night_pid_mean, na.rm = TRUE),
      ln_within_10  = late_night_within / 10,
      ln_between_10 = late_night_between / 10
    ) |>
    filter(!is.na(sleep_quality_ord))
  d
}

diary_imputed_prep_h2 <- function(d) {
  d <- diary_imputed_prep_h1(d)
  d |> filter(!is.na(chronotype_centered))
}

# H1 formula (same as complete-case)
diary_h1_formula_imp <- sleep_quality_ord ~
  ln_within_10 + ln_between_10 +
  age_scaled + bmi_scaled + SES_index_scaled + is_weekend + (1 | pid)

# H2 formula (same as complete-case)
diary_h2_formula_imp <- sleep_quality_ord ~
  ln_within_10 * chronotype_centered +
  ln_between_10 * chronotype_centered +
  age_scaled + bmi_scaled + SES_index_scaled + is_weekend + (1 | pid)

message("Fitting diary H1 on imputed data (Rubin's rules)...")
diary_h1_rubin <- fit_on_imputations(
  diary_imputed_raw,
  prep_fn = diary_imputed_prep_h1,
  fit_fn = function(d) clmm(
    diary_h1_formula_imp, data = d, link = "probit", Hess = TRUE,
    control = clmm.control(maxIter = 50000, gradTol = 0.5)
  )
)
saveRDS(diary_h1_rubin, "output/models/imputed_diary_h1_rubin.rds")

message("Fitting diary H2 on imputed data (Rubin's rules)...")
diary_h2_rubin <- fit_on_imputations(
  diary_imputed_raw,
  prep_fn = diary_imputed_prep_h2,
  fit_fn = function(d) clmm(
    diary_h2_formula_imp, data = d, link = "probit", Hess = TRUE,
    control = clmm.control(maxIter = 50000, gradTol = 0.5)
  )
)
saveRDS(diary_h2_rubin, "output/models/imputed_diary_h2_rubin.rds")

message("Diary imputed models saved.")
```

```{r}
#| label: diary-imputed-load
#| echo: false
#| message: false
#| warning: false
#| cache: false

# Load diary Rubin bundles from disk if not just fitted
if (!exists("diary_h1_rubin") || is.null(diary_h1_rubin)) {
  diary_h1_rubin <- tryCatch(
    readRDS("output/models/imputed_diary_h1_rubin.rds"),
    error = function(e) NULL
  )
}
if (!exists("diary_h2_rubin") || is.null(diary_h2_rubin)) {
  diary_h2_rubin <- tryCatch(
    readRDS("output/models/imputed_diary_h2_rubin.rds"),
    error = function(e) NULL
  )
}
```

```{r}
#| label: diary-main-effects
#| echo: false
#| message: false
#| warning: false
#| cache: false

# Complete-case model paths (fallback)
diary_h1_path <- "output/models/diary_model_h1_sleepquality.rds"
diary_h2_path <- "output/models/diary_model_h2_sleepquality.rds"

# H1 effects: use pooled Rubin estimates if available, else fall back to complete-case
if (!is.null(diary_h1_rubin)) {
  diary_h1_within <- get_pooled_effect(diary_h1_rubin, "ln_within_10", exponentiate = FALSE)
  diary_h1_between <- get_pooled_effect(diary_h1_rubin, "ln_between_10", exponentiate = FALSE)
} else {
  diary_h1_within <- get_h2_interaction(diary_h1_path, "ln_within_10", exponentiate = FALSE)
  diary_h1_between <- get_h2_interaction(diary_h1_path, "ln_between_10", exponentiate = FALSE)
}

# H2 effects: use pooled Rubin estimates if available, else fall back to complete-case
if (!is.null(diary_h2_rubin)) {
  diary_h2_within_chrono <- get_pooled_effect(diary_h2_rubin, "ln_within_10:chronotype_centered",
                                               exponentiate = FALSE)
  diary_h2_between_chrono <- get_pooled_effect(diary_h2_rubin, "chronotype_centered:ln_between_10",
                                                exponentiate = FALSE)
  diary_h2_chrono_main <- get_pooled_effect(diary_h2_rubin, "chronotype_centered",
                                             exponentiate = FALSE)
} else {
  diary_h2_within_chrono <- get_h2_interaction(diary_h2_path,
    "ln_within_10:chronotype_centered", exponentiate = FALSE)
  diary_h2_between_chrono <- get_h2_interaction(diary_h2_path,
    "chronotype_centered:ln_between_10", exponentiate = FALSE)
  diary_h2_chrono_main <- get_h2_interaction(diary_h2_path,
    "chronotype_centered", exponentiate = FALSE)
}

# Spline comparison info
diary_spline_compare_loaded <- if (file.exists("output/models/diary_spline_compare.rds")) {
  readRDS("output/models/diary_spline_compare.rds")
} else {
  NULL
}

# Extract raw numeric values for inline interpretation (sig tests, direction)
extract_pooled_est <- function(bundle, term) {
  if (is.null(bundle) || is.null(bundle$pooled)) return(list(est = NA, ci_low = NA, ci_high = NA, p = NA))
  s <- summary(bundle$pooled, conf.int = TRUE)
  row <- s[s$term == term, , drop = FALSE]
  if (nrow(row) == 0) return(list(est = NA, ci_low = NA, ci_high = NA, p = NA))
  list(est = row$estimate, ci_low = row$`2.5 %`, ci_high = row$`97.5 %`, p = row$p.value)
}

extract_clmm_est <- function(path, term, exponentiate = FALSE) {
  if (!file.exists(path)) return(list(est = NA, ci_low = NA, ci_high = NA, p = NA))
  m <- readRDS(path)
  ct <- as.data.frame(coef(summary(m)))
  ct$Parameter <- rownames(ct)
  row <- ct[ct$Parameter == term, , drop = FALSE]
  if (nrow(row) == 0) return(list(est = NA, ci_low = NA, ci_high = NA, p = NA))
  est <- as.numeric(row[["Estimate"]])
  se  <- as.numeric(row[["Std. Error"]])
  p   <- suppressWarnings(as.numeric(row[["Pr(>|z|)"]]))
  crit <- qnorm(0.975)
  ci_low <- est - crit * se; ci_high <- est + crit * se
  if (exponentiate) { est <- exp(est); ci_low <- exp(ci_low); ci_high <- exp(ci_high) }
  list(est = est, ci_low = ci_low, ci_high = ci_high, p = p)
}

if (!is.null(diary_h1_rubin)) {
  diary_h1_w  <- extract_pooled_est(diary_h1_rubin, "ln_within_10")
  diary_h1_b  <- extract_pooled_est(diary_h1_rubin, "ln_between_10")
} else {
  diary_h1_w  <- extract_clmm_est(diary_h1_path, "ln_within_10")
  diary_h1_b  <- extract_clmm_est(diary_h1_path, "ln_between_10")
}

if (!is.null(diary_h2_rubin)) {
  diary_h2_wc <- extract_pooled_est(diary_h2_rubin, "ln_within_10:chronotype_centered")
  diary_h2_bc <- extract_pooled_est(diary_h2_rubin, "chronotype_centered:ln_between_10")
  diary_h2_cm <- extract_pooled_est(diary_h2_rubin, "chronotype_centered")
} else {
  diary_h2_wc <- extract_clmm_est(diary_h2_path, "ln_within_10:chronotype_centered")
  diary_h2_bc <- extract_clmm_est(diary_h2_path, "chronotype_centered:ln_between_10")
  diary_h2_cm <- extract_clmm_est(diary_h2_path, "chronotype_centered")
}

# Helper for significance description
sig_desc <- function(p) {
  if (is.na(p)) return("not statistically significant")
  if (p < .001) "statistically significant"
  else if (p < .05) "statistically significant"
  else if (p < .10) "marginally significant"
  else "not statistically significant"
}

# Marginal probability computation for diary H1 (probit CLMM, between-person)
# Use pooled estimates (Rubin's rules) when available, else single complete-case fit
diary_h1_marginal <- if (!is.null(diary_h1_rubin)) {
  tryCatch({
    pv        <- pool_vcov(diary_h1_rubin$fits)
    b_between <- pv$qbar["ln_between_10"]
    t_fair_poor <- average_re_quantity(diary_h1_rubin$fits, function(f) f$alpha["Fair|Poor"])
    sigma_u   <- average_re_quantity(diary_h1_rubin$fits, function(f) sqrt(f$ST$pid[1, 1]))
    marginal_sd <- sqrt(1 + sigma_u^2)
    pp_per_10min <- dnorm((b_between * 0 - t_fair_poor) / marginal_sd) * b_between / marginal_sd * 100
    list(pp_per_10min = pp_per_10min)
  }, error = function(e) NULL)
} else {
  tryCatch({
    m <- readRDS(diary_h1_path)
    b_between <- m$beta["ln_between_10"]
    t_fair_poor <- m$alpha["Fair|Poor"]
    sd_re <- m$ST$pid[1, 1]
    marginal_sd <- sqrt(1 + sd_re^2)
    pp_per_10min <- dnorm((b_between * 0 - t_fair_poor) / marginal_sd) * b_between / marginal_sd * 100
    list(pp_per_10min = pp_per_10min)
  }, error = function(e) NULL)
}
```

To complement the panel-level analyses, we examined daily diary reports of sleep quality as a function of late-night gaming in the preceding 24 hours. Sleep quality was recorded on a 5-level ordinal scale (Very poor, Poor, Fair, Good, Very good) in each diary entry. We fitted cumulative link mixed models (CLMM; `ordinal::clmm`) with random intercepts for participants to account for repeated measurements. To separate within-person fluctuations from between-person differences in late-night gaming, we decomposed the trimmed late-night gaming variable (capped at the 99th percentile) into a within-person component (daily deviation from the participant's own mean) and a between-person component (participant mean centered at the grand mean). The analytical diary sample comprised `r diary_n_participants` participants contributing `r format(diary_n_entries, big.mark = ",")` diary entries with valid sleep quality ratings. Covariates included age, BMI, SES (all scaled), and a weekend indicator. Region is excluded from both diary models because all diary participants are US-based. These diary analyses are exploratory and were not formally pre-registered. Gender was initially considered as a random intercept but the random intercept for gender with only three levels did not converge; gender is therefore excluded from all diary models. A probit link was used (rather than logit) to ensure reliable convergence for both models. Late-night gaming is expressed in per-10-minute increments (i.e., the effect of each additional 10 minutes of late-night gaming) to aid interpretability while maintaining numerical stability. In the ordered factor coding, higher categories correspond to worse sleep quality (5 = "Very poor"), so positive coefficients indicate a higher probability of worse sleep. Covariates (age, BMI, SES) are scaled within the diary subsample. Missing SES values for diary participants with unrecognised employment categories (n = 23) were imputed via MICE (PMM, m = 5) using age, BMI, gender, and region as predictors, consistent with the panel-level missing-data strategy.

For the direct-effects model (H1), the within-person effect of late-night gaming on sleep quality was `r diary_h1_within` (probit coefficient per 10 additional minutes). On days when participants gamed more than their own average, the association with sleep quality was `r ifelse(!is.na(diary_h1_w$p), sig_desc(diary_h1_w$p), "significance not available")`. The between-person effect was `r diary_h1_between`, indicating that participants who habitually engaged in more late-night gaming tended to report `r ifelse(!is.na(diary_h1_b$est) && diary_h1_b$est > 0, "worse", "better")` sleep quality on average; this between-person effect was `r ifelse(!is.na(diary_h1_b$p), sig_desc(diary_h1_b$p), "of uncertain significance")`. On the probability scale, each additional 10 minutes of habitual daily late-night gaming is associated with an approximate `r if (!is.null(diary_h1_marginal)) sprintf("%.1f", diary_h1_marginal$pp_per_10min)` percentage-point increase in the probability of reporting poor or very poor sleep, broadly consistent with the panel-level H1a estimate (`r if (!is.null(h1a_marginal)) sprintf("%.1f", h1a_marginal$pp_per_10min)` pp per 10 min).
`r if (!is.null(diary_spline_compare_loaded)) paste0("We compared the linear specification against natural spline alternatives (df = 2-6); the best-fitting spline (df = ", diary_spline_compare_loaded$df[1], ") had a \\u0394AIC of ", sprintf("%.1f", diary_spline_compare_loaded$AIC[1] - diary_spline_compare_loaded$linear_AIC[1]), " relative to the linear model, suggesting that the linear parameterization provided an adequate fit.")`
Diary subsample characteristics are presented in @tbl-diary-demographics in the Appendix.

For the chronotype moderation model (H2), the interaction between within-person late-night gaming and chronotype was `r diary_h2_within_chrono`, and the interaction between between-person late-night gaming and chronotype was `r diary_h2_between_chrono`. The within-person interaction was `r ifelse(!is.na(diary_h2_wc$p), sig_desc(diary_h2_wc$p), "of uncertain significance")`, indicating that the day-to-day link between late-night gaming and sleep quality was `r ifelse(!is.na(diary_h2_wc$est) && diary_h2_wc$est > 0, "somewhat stronger for evening types (positive interaction)", "somewhat weaker for evening types")`. The between-person interaction was `r ifelse(!is.na(diary_h2_bc$p), sig_desc(diary_h2_bc$p), "of uncertain significance")`, suggesting that the trait-level association between habitual late-night gaming and sleep quality did not meaningfully vary across chronotypes. The main effect of chronotype was `r diary_h2_chrono_main`; a later chronotype was associated with `r ifelse(!is.na(diary_h2_cm$est) && diary_h2_cm$est > 0, "worse", "better")` sleep quality (this effect was `r ifelse(!is.na(diary_h2_cm$p), sig_desc(diary_h2_cm$p), "of uncertain significance")`). These diary-based findings are exploratory and should be interpreted with caution.

```{r}
#| label: fig-diary-sleepquality
#| fig-cap: "Predicted probability of poor sleep quality as a function of average late-night gaming (diary study, between-person effect). The line shows the marginal probability of reporting poor sleep quality (Poor or Very poor on the 5-level ordinal scale) across the range of participants' average daily late-night gaming, centered at the grand mean. The ribbon represents 95% confidence intervals derived from the between-person gaming coefficient SE. Vertical reference lines mark the median (dotted), mean (dashed), 75th percentile (dotted), and 90th percentile (dashed) of the sample distribution. The marginal density plot (top panel) shows the distribution of between-person late-night gaming. Predictions are population-average probabilities (marginalised over the participant random intercept) derived from the diary H1 probit CLMM, with the within-person component held at zero and other covariates at their reference/mean values."
#| fig-width: 9
#| fig-height: 6
#| code-summary: "Diary: probability of poor sleep quality by between-person late-night gaming"
#| echo: false
#| message: false
#| warning: false

diary_h1_model_available <- !is.null(diary_h1_rubin) || file.exists("output/models/diary_model_h1_sleepquality.rds")

if (diary_h1_model_available) {
  library(ordinal)
  library(patchwork)

  # Prediction grid (per-10-min units: -6 to 9 = -60 to 90 min)
  ln_seq <- seq(-6, 9, by = 0.1)

  # MARGINAL exceedance probability — same method as panel Figure 2.
  # For pooled (Rubin's rules): use pool_vcov() for b and SE, average_re_quantity()
  # for alpha_j and sigma_u across imputations.
  # For complete-case fallback: load single model directly.
  if (!is.null(diary_h1_rubin)) {
    pv_d        <- pool_vcov(diary_h1_rubin$fits)
    b_between   <- pv_d$qbar["ln_between_10"]
    se_between  <- sqrt(pv_d$vcov["ln_between_10", "ln_between_10"])
    alpha_j_d   <- average_re_quantity(diary_h1_rubin$fits, function(f) f$alpha["Fair|Poor"])
    sigma_u_d   <- average_re_quantity(diary_h1_rubin$fits, function(f) sqrt(f$ST$pid[1, 1]))
  } else {
    diary_h1_fig <- readRDS("output/models/diary_model_h1_sleepquality.rds")
    sigma_u_d   <- sqrt(diary_h1_fig$ST$pid[1, 1])
    alpha_j_d   <- diary_h1_fig$alpha["Fair|Poor"]
    b_between   <- coef(diary_h1_fig)["ln_between_10"]
    se_between  <- sqrt(vcov(diary_h1_fig)["ln_between_10", "ln_between_10"])
  }

  denom_d     <- sqrt(1 + sigma_u_d^2)

  lp_d    <- b_between * ln_seq
  lp_d_lo <- (b_between - 1.96 * se_between) * ln_seq
  lp_d_hi <- (b_between + 1.96 * se_between) * ln_seq

  df_between <- tibble(
    gaming_min = ln_seq * 10,
    exc.prob  = pnorm((alpha_j_d - lp_d)    / denom_d, lower.tail = FALSE),
    asymp.LCL = pnorm((alpha_j_d - lp_d_lo) / denom_d, lower.tail = FALSE),
    asymp.UCL = pnorm((alpha_j_d - lp_d_hi) / denom_d, lower.tail = FALSE)
  )

  x_range  <- c(-60, 90)
  x_breaks <- seq(-60, 90, by = 30)

  between_stats <- list(
    mean   = mean(diary_ocat$late_night_between, na.rm = TRUE),
    median = median(diary_ocat$late_night_between, na.rm = TRUE),
    p75    = quantile(diary_ocat$late_night_between, 0.75, na.rm = TRUE),
    p90    = quantile(diary_ocat$late_night_between, 0.90, na.rm = TRUE)
  )

  p_dist <- ggplot(
    data.frame(x = diary_ocat$late_night_between[!is.na(diary_ocat$late_night_between)]),
    aes(x = x)
  ) +
    geom_density(fill = "gray70", color = "gray50", alpha = 0.6) +
    geom_vline(xintercept = between_stats$median, linetype = "dotted", color = "gray40", linewidth = 0.5) +
    geom_vline(xintercept = between_stats$mean,   linetype = "dashed",  color = "gray40", linewidth = 0.5) +
    geom_vline(xintercept = between_stats$p75,    linetype = "dotted", color = "gray40", linewidth = 0.5) +
    geom_vline(xintercept = between_stats$p90,    linetype = "dashed",  color = "gray40", linewidth = 0.5) +
    scale_x_continuous(breaks = x_breaks, labels = paste0(x_breaks, " min"), limits = x_range) +
    labs(y = "Density", x = NULL) +
    theme_minimal(base_size = 11) +
    theme(
      panel.grid.minor = element_blank(),
      axis.text.x      = element_blank(),
      axis.title.x     = element_blank(),
      plot.margin      = margin(5, 5, 0, 5)
    )

  y_top <- max(df_between$asymp.UCL, na.rm = TRUE)
  y_ann <- y_top * 0.88

  p_main <- ggplot(df_between, aes(x = gaming_min, y = exc.prob)) +
    geom_vline(xintercept = between_stats$median, linetype = "dotted", color = "gray40", linewidth = 0.5) +
    geom_vline(xintercept = between_stats$mean,   linetype = "dashed",  color = "gray40", linewidth = 0.5) +
    geom_vline(xintercept = between_stats$p75,    linetype = "dotted", color = "gray40", linewidth = 0.5) +
    geom_vline(xintercept = between_stats$p90,    linetype = "dashed",  color = "gray40", linewidth = 0.5) +
    annotate("text", x = between_stats$median, y = y_ann, label = "Median",
             angle = 90, vjust = -0.5, size = 3, color = "gray40") +
    annotate("text", x = between_stats$mean,   y = y_ann, label = "Mean",
             angle = 90, vjust = -0.5, size = 3, color = "gray40") +
    annotate("text", x = between_stats$p75,    y = y_ann, label = "75th %ile",
             angle = 90, vjust = -0.5, size = 3, color = "gray40") +
    annotate("text", x = between_stats$p90,    y = y_ann, label = "90th %ile",
             angle = 90, vjust = -0.5, size = 3, color = "gray40") +
    geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL),
                alpha = 0.4, fill = "#f46d43", color = "#f46d43",
                linewidth = 0.3, linetype = "dashed") +
    geom_line(linewidth = 1.5, color = "#d62828") +
    scale_x_continuous(breaks = x_breaks, labels = paste0(x_breaks, " min"), limits = x_range) +
    scale_y_continuous(labels = scales::percent_format(), limits = c(0, y_top * 1.2)) +
    labs(
      x = "Average Daily Late-Night Gaming, Centered (min/day)",
      y = "Predicted Probability of Poor Sleep Quality"
    ) +
    theme_minimal(base_size = 11) +
    theme(panel.grid.minor = element_blank(), plot.margin = margin(0, 5, 5, 5))

  (p_dist / p_main) +
    plot_layout(heights = c(1, 3)) +
    plot_annotation(
      title = "Diary: Probability of Poor Sleep Quality by Late-Night Gaming"
    )
} else {
  message("Diary H1 model not found. Run with eval_diary = TRUE or eval_diary_imputed = TRUE first.")
}
```

```{r}
#| label: tbl-diary-h1h2
#| tbl-cap: "Diary CLMM Regression Results: H1 (Direct Effects) and H2 (Chronotype Moderation) for Sleep Quality"
#| echo: false
#| message: false
#| warning: false

library(tinytable)
library(modelsummary)

# Use imputed (Rubin's rules) models when available, else fall back to complete-case
diary_use_imputed <- !is.null(diary_h1_rubin) && !is.null(diary_h2_rubin)

if (diary_use_imputed) {
  diary_models_combined <- list(
    "H1: Sleep Quality" = diary_h1_rubin$pooled,
    "H2: Chronotype Moderation" = diary_h2_rubin$pooled
  )
  # Coef names from pooled summary
  all_coef_names <- unique(c(
    summary(diary_h1_rubin$pooled)$term,
    summary(diary_h2_rubin$pooled)$term
  ))
  diary_add_rows <- get_custom_rows_pooled(
    list("H1: Sleep Quality" = diary_h1_rubin, "H2: Chronotype Moderation" = diary_h2_rubin)
  )
  tbl_note_source <- "Estimates pooled across 20 multiply imputed datasets using Rubin's rules."
} else {
  diary_model_h1_loaded <- tryCatch(readRDS("output/models/diary_model_h1_sleepquality.rds"), error = function(e) NULL)
  diary_model_h2_loaded <- tryCatch(readRDS("output/models/diary_model_h2_sleepquality.rds"), error = function(e) NULL)
  diary_models_combined <- list(
    "H1: Sleep Quality" = diary_model_h1_loaded,
    "H2: Chronotype Moderation" = diary_model_h2_loaded
  )
  all_coef_names <- if (!is.null(diary_model_h1_loaded) && !is.null(diary_model_h2_loaded)) {
    unique(c(names(coef(diary_model_h1_loaded)), names(coef(diary_model_h2_loaded))))
  } else character(0)
  diary_add_rows <- get_custom_rows(Filter(Negate(is.null), diary_models_combined))
  tbl_note_source <- "Complete-case estimates (imputed diary models not found)."
}

diary_any_model <- !is.null(diary_models_combined[["H1: Sleep Quality"]]) &&
                   !is.null(diary_models_combined[["H2: Chronotype Moderation"]])

if (diary_any_model) {
  diary_coef_map <- c(
    "ln_within_10"                       = "LN gaming within-person (per 10 min)",
    "ln_between_10"                      = "LN gaming between-person (per 10 min)",
    "chronotype_centered"                = "Chronotype (h, centered)",
    "ln_within_10:chronotype_centered"   = "LN within × Chronotype",
    "chronotype_centered:ln_between_10"  = "LN between × Chronotype",
    "age_scaled"                         = "Age (scaled)",
    "bmi_scaled"                         = "BMI (scaled)",
    "SES_index_scaled"                   = "SES (scaled)"
  )

  weekend_map <- c()
  if ("is_weekendTRUE" %in% all_coef_names) {
    weekend_map <- c("is_weekendTRUE" = "Day: Weekend")
  } else if ("is_weekend" %in% all_coef_names) {
    weekend_map <- c("is_weekend" = "Day: Weekend")
  }
  diary_coef_map <- c(diary_coef_map, weekend_map)

  tt_diary <- modelsummary(
    diary_models_combined,
    fmt = fmt_sprintf("%.2f"),
    estimate = "{estimate} [{conf.low}, {conf.high}]{stars}",
    statistic = NULL,
    gof_omit = "AIC|BIC|Log|RMSE|R2|ICC|Num|nobs",
    coef_map = diary_coef_map,
    add_rows = diary_add_rows,
    notes = c(
      "+ p < 0.10, * p < 0.05, ** p < 0.01, *** p < 0.001",
      "LN = late-night. Confidence intervals shown in brackets.",
      tbl_note_source,
      "Cumulative link mixed models (random intercept for participant) on 5-level ordinal sleep quality (positive coefficients = higher probability of worse sleep). Both H1 and H2 use probit link. Late-night gaming expressed per 10 minutes; chronotype in centered hours; age, BMI, SES scaled within the diary subsample. Region excluded: all diary participants are US-only.",
      "Within = daily deviation from person mean; Between = person mean - grand mean."
    ),
    output = "tinytable"
  )
  dir.create("output/tables", showWarnings = FALSE, recursive = TRUE)
  tryCatch(
    save_tt(tt_diary, "output/tables/tbl-diary-h1h2.png", overwrite = TRUE),
    error = function(e) message("Note: PNG export skipped (", e$message, ")")
  )
  tt_diary
} else {
  cat("Diary models not found. Render with `-P refit_diary_imputed:true` to generate them.")
}
```

# Discussion

Our findings from a sample of adult gamers revealed patterns consistent with prior research: while sleep duration remained relatively preserved (7.2 hours), sleep quality showed greater impairment, with 64% classified as poor sleepers. These patterns suggest that sleep quality and architecture may be more vulnerable to disruption than sleep duration in the context of regular gaming behavior.

## Limitations

First, our telemetry and survey instruments capture console and PC play but exclude mobile platforms. Gaming on smartphones and tablets forms a substantial share of late-night leisure, particularly in bed where devices are at arm’s reach and can bypass household curfews. Phones also double as tools for emotional regulation, relaxation, and boredom relief before sleep or when people wake up during the night, further increasing the odds that meaningful late-night gaming takes place on devices we do not observe. Any late-night gaming minutes that shifted to mobile therefore go unmeasured, likely biasing exposure downward and attenuating observed associations with sleep. Related activities on the same devices like short-form video viewing, doomscrolling, messaging are also missing, preventing us from disentangling gaming-specific effects from broader nocturnal screen engagement.

Second, the analytic sample consists of early adults. This demographic is commercially important for platform gaming, yet the educational and developmental stakes that motivate policy debates revolve around children and adolescents. Younger players have different sleep physiology, earlier school start times, tighter parental controls, and distinct motivational profiles that could amplify or dampen the impact of late-night play. Our estimates should not be generalized to pediatric populations (or to older adults with different health burdens) without direct evidence.

Third, recruitment relied on UK- and US-based majority Prolific panels, which may diverge markedly from the general gaming population for several reasons. Prolific users are self-selected online workers with reliable broadband, high digital literacy, and the patience for repeated surveys. Platform eligibility emphasizes UK/US residency, English fluency, and adherence to Prolific’s compliance checks, which excludes players in other linguistic, cultural, and regulatory contexts. Incentives attract habitual survey takers with flexible schedules, potentially underrepresenting shift workers, caregivers, or gamers with highly variable sleep routines. Community standards and attention checks may dissuade individuals engaged in more extreme, unregulated, or stigmatized gaming profiles from participating, while payment infrastructures and screening filters can limit participation from lower-income players, people without stable online payment methods, and those who distrust research platforms.

Finally, although logged telemetry is widely recommended for media-effects research—and we did find that it is strongly correlated with self-reported video game play—it is not without limitations. Session-level records cannot distinguish between active engagement and idle time, and this ambiguity is especially pronounced for late-night gaming, where consoles or PCs might stay on while players drift off to sleep or attend to household chores. Such overestimation of exposure could mask true associations or produce spurious late-night “activity” that does not reflect cognitive arousal.

Taken together, the lack of mobile telemetry, the age range of the sample, the geographic/panel constraints, and the inability to rule out idle device time mean that our findings speak most directly to digitally engaged young adults in the UK and US who play on major console and PC ecosystems. Extending the conclusions to other regions, life stages, device ecologies, or logging systems should await complementary data sources.

# References

:::{#refs}
:::

```{=typst}
#show: appendix.with()
```
# Appendix: Complete-Case Regression Tables

The tables below reproduce the H1 and H2 regression summaries using complete-case data (excluding observations where the outcome was imputed). They correspond to the complete-case analyses referenced in the main text and serve as a sensitivity check alongside the imputed results.

### H1 Sensitivity Analysis: Complete-Case (Non-Imputed Data)

These models use only observations where the outcome variable was NOT imputed (complete-case analysis per outcome). Sample sizes are smaller than the main imputed analyses, as rows with imputed outcomes are excluded.

```{r}
#| label: appendix-h1-original
#| echo: false
#| output: asis

if ("completecase" %in% names(h1_models)) {
  tt_h1_cc <- modelsummary(
    h1_models[["completecase"]],
    fmt = fmt_sprintf("%.2f"),
    estimate  = "{estimate} [{conf.low}, {conf.high}]{stars}",
    statistic = NULL,
    gof_omit = "AIC|BIC|Log|RMSE|R2|ICC|Num|nobs",
    coef_map = coef_map_h1,
    add_rows = get_custom_rows(h1_models[["completecase"]]),
    notes = c("+ p < 0.10, * p < 0.05, ** p < 0.01, *** p < 0.001",
              "LN = late-night. Confidence intervals shown in brackets.",
              "ICC = Intraclass Correlation Coefficient (adjusted)."),
    title = "Appendix H1: Effects of Late-Night Gaming (Complete-Case Analysis)"
  )
  dir.create("output/tables", showWarnings = FALSE, recursive = TRUE)
  tryCatch(
    save_tt(tt_h1_cc, "output/tables/appendix-h1-completecase.png", overwrite = TRUE),
    error = function(e) message("Note: PNG export skipped (", e$message, ")")
  )
  tt_h1_cc
} else {
  cat("Complete-case H1 models are unavailable. Render with `-P refit_appendix:true` to regenerate them.\n")
}
```

```{=typst}
#pagebreak()
```


### H2 Sensitivity Analysis: Complete-Case (Non-Imputed Data)

These models use only observations where the outcome variable was NOT imputed (complete-case analysis per outcome). Sample sizes are smaller than the main imputed analyses, as rows with imputed outcomes are excluded.

```{r}
#| label: appendix-h2-original
#| echo: false
#| output: asis

if ("completecase" %in% names(h2_models)) {
  tt_h2_cc <- modelsummary(
    h2_models[["completecase"]],
    fmt = fmt_sprintf("%.2f"),
    estimate  = "{estimate} [{conf.low}, {conf.high}]{stars}",
    statistic = NULL,
    gof_omit = "AIC|BIC|Log|RMSE|R2|ICC|Num|nobs",
    coef_map = coef_map_h2,
    add_rows = get_custom_rows(h2_models[["completecase"]]),
    notes = c("+ p < 0.10, * p < 0.05, ** p < 0.01, *** p < 0.001",
              "LN = late-night. Confidence intervals shown in brackets.",
              "ICC = Intraclass Correlation Coefficient (adjusted)."),
    title = "Appendix H2: Chronotype Moderation (Complete-Case Analysis)"
  )
  dir.create("output/tables", showWarnings = FALSE, recursive = TRUE)
  tryCatch(
    save_tt(tt_h2_cc, "output/tables/appendix-h2-completecase.png", overwrite = TRUE),
    error = function(e) message("Note: PNG export skipped (", e$message, ")")
  )
  tt_h2_cc
} else {
  cat("Complete-case H2 models are unavailable. Render with `-P refit_appendix:true` to regenerate them.\n")
}
```

```{=typst}
#pagebreak()
```

### Wave-Level Missingness (Pre-Imputation Data)

We summarised the extent of missingness for the key self-report outcomes across each survey wave in the raw data before imputation. The `Observations` column reports the number of participants in a wave who completed at least one of the listed measures; the percentages in each row are calculated relative to that wave-specific participant count.

```{r}
#| label: appendix-wave-missingness
#| echo: false
#| tbl-cap: "Wave-level missingness for key self-report measures (pre-imputation data)"
#| message: false
#| warning: false

library(tinytable)

missing_measures <- c("psqi_comp1_quality", "total_hours_sleep", "epsTotal", "wemwbs")
missing_flags <- paste0(missing_measures, "_imputed_flag")

wave_observations <- selfreport |>
  filter(!is.na(wave)) |>
  mutate(
    has_any_measure = if_any(all_of(missing_measures), ~!is.na(.x))
  ) |>
  group_by(wave) |>
  summarise(
    Observations = dplyr::n_distinct(pid[has_any_measure]),
    .groups = "drop"
  )

wave_missing_counts <- selfreport |>
  filter(!is.na(wave)) |>
  group_by(wave) |>
  summarise(
    across(all_of(missing_flags), ~sum(.x, na.rm = TRUE)),
    .groups = "drop"
  ) |>
  left_join(wave_observations, by = "wave") |>
  arrange(wave) |>
  mutate(
    Observations = dplyr::coalesce(Observations, 0L),
    Wave = as.integer(wave)
  ) |>
  select(Wave, Observations, all_of(missing_flags)) |>
  mutate(
    across(
      all_of(missing_flags),
      ~if_else(
        Observations > 0,
        sprintf("%d (%.1f%%)", .x, 100 * .x / Observations),
        sprintf("%d (--)", .x)
      )
    )
  )

wave_missing_table <- wave_missing_counts |>
  rename(
    `Sleep quality (PSQI item 6)` = psqi_comp1_quality_imputed_flag,
    `Sleep duration (hours)` = total_hours_sleep_imputed_flag,
    `Daytime sleepiness (ESS)` = epsTotal_imputed_flag,
    `Wellbeing (SWEMWBS)` = wemwbs_imputed_flag
  ) |>
  mutate(
    across(
      c(`Sleep quality (PSQI item 6)`, `Sleep duration (hours)`, `Daytime sleepiness (ESS)`),
      ~if_else(Wave %in% c(1L, 3L, 5L), "", .x)
    )
  )

tinytable::tt(
  wave_missing_table,
  caption = "Wave-level missingness for key self-report measures (pre-imputation data). Entries show the number and percentage of participants missing each measure within a wave."
)
```

```{=typst}
#pagebreak()
```

# Appendix: PSQI Global Score Sensitivity Analysis

This section presents a sensitivity analysis using the PSQI global score (psqi_global) as an alternative sleep quality outcome. The PSQI global score is the sum of all 7 PSQI component scores (range 0-21, higher = worse sleep quality), providing a continuous measure compared to the ordinal item 6 (psqi_comp1_quality) used in the pre-registered H1a hypothesis.

We fit linear mixed-effects models with the same predictor structure as H1d (daytime sleepiness), comparing results for imputed data vs. complete-case data (excluding observations where PSQI global was imputed).

```{r}
#| label: tbl-appendix-psqi-global
#| echo: false
#| tbl-cap: "Sensitivity Analysis: PSQI Global Score Models (Imputed vs. Original)"
#| message: false
#| warning: false

# Prepare complete-case monthly gaming data with psqi_global
gamingMonthly_completecase <- gaming_inputs$completecase$gamingMonthly |>
  enforce_monthly_wave_subset() |>
  filter(!is.na(psqi_global))

# --- Imputed: Rubin's rules ---
panel_psqi_prep_monthly <- function(d) {
  inputs <- build_gaming_inputs(d)
  inputs$gamingMonthly |>
    enforce_monthly_wave_subset() |>
    filter(!is.na(psqi_global))
}

panel_psqi_prep_monthly_chrono <- function(d) {
  inputs <- build_gaming_inputs(d)
  inputs$gamingMonthly |>
    enforce_monthly_wave_subset() |>
    filter(!is.na(psqi_global), !is.na(msf_sc_centered))
}

panel_psqi_rubin_base <- fit_on_imputations(
  imputed_long,
  prep_fn = panel_psqi_prep_monthly,
  fit_fn = function(d) lmer(
    psqi_global ~ ln_monthly_avg_minutes_played_10 +
      age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend +
      (1 | pid) + (1 | gender),
    data = d, control = lmerControl(optimizer = "bobyqa")
  )
)
model_psqi_global_imputed <- panel_psqi_rubin_base$pooled

panel_psqi_rubin_chrono <- fit_on_imputations(
  imputed_long,
  prep_fn = panel_psqi_prep_monthly_chrono,
  fit_fn = function(d) lmer(
    psqi_global ~ ln_monthly_avg_minutes_played_10 * msf_sc_centered +
      age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend +
      (1 | pid) + (1 | gender),
    data = d, control = lmerControl(optimizer = "bobyqa")
  )
)
model_psqi_global_imputed_chrono <- panel_psqi_rubin_chrono$pooled

# --- Complete-case: single fit (unchanged) ---
model_psqi_global_completecase <- lmer(
  psqi_global ~ ln_monthly_avg_minutes_played_10 +
    age_scaled + bmi_scaled + SES_index_scaled + region  + isWeekend +
    (1 | pid) + (1 | gender),
  data = gamingMonthly_completecase,
  control = lmerControl(optimizer = "bobyqa")
)

model_psqi_global_completecase_chrono <- lmer(
  psqi_global ~ ln_monthly_avg_minutes_played_10 * msf_sc_centered +
    age_scaled + bmi_scaled + SES_index_scaled + region  + isWeekend +
    (1 | pid) + (1 | gender),
  data = gamingMonthly_completecase,
  control = lmerControl(optimizer = "bobyqa")
)

# Create comparison table
# Match the row structure of the H1/H2 summaries for easier cross-referencing
psqi_label_lookup <- c(
  "ln_monthly_avg_minutes_played_10" = "Daily LN gaming (per 10 min/day, monthly)",
  "age_scaled" = "Age (scaled)",
  "bmi_scaled" = "BMI (scaled)",
  "SES_index_scaled" = "SES (scaled)",
  "msf_sc_centered" = "Chronotype (h, centered)",
  "ln_monthly_avg_minutes_played_10:msf_sc_centered" = "LN gaming × Chronotype (h, monthly)",
  "isWeekend" = "Day: Weekend",
  "SD (Intercept pid)" = "SD (Intercept | Participant)",
  "SD (Intercept gender)" = "SD (Intercept | Gender)",
  "SD (Observations)" = "SD (Residual)"
)

psqi_base_terms <- c(
  "ln_monthly_avg_minutes_played_10",
  "age_scaled",
  "bmi_scaled",
  "SES_index_scaled"
)

psqi_moderator_terms <- c(
  "msf_sc_centered",
  "ln_monthly_avg_minutes_played_10:msf_sc_centered"
)

psqi_random_terms <- c(
  "SD (Intercept pid)",
  "SD (Intercept gender)",
  "SD (Observations)"
)

psqi_base_map <- psqi_label_lookup[psqi_base_terms]
psqi_base_map <- psqi_base_map[!is.na(psqi_base_map)]

psqi_moderator_map <- psqi_label_lookup[psqi_moderator_terms]
psqi_moderator_map <- psqi_moderator_map[!is.na(psqi_moderator_map)]

psqi_random_map <- psqi_label_lookup[psqi_random_terms]
psqi_random_map <- psqi_random_map[!is.na(psqi_random_map)]

# Extract fixed effect names from all models (pooled mipo or lmerMod)
.get_term_names <- function(m) {
 if (inherits(m, "mipo")) {
    summary(m)$term
  } else {
    names(lme4::fixef(m))
  }
}
fixef_names <- unique(c(
  .get_term_names(model_psqi_global_imputed),
  .get_term_names(model_psqi_global_completecase),
  .get_term_names(model_psqi_global_imputed_chrono),
  .get_term_names(model_psqi_global_completecase_chrono)
))
fixef_names <- fixef_names[fixef_names != "(Intercept)"]

region_terms <- fixef_names[str_detect(fixef_names, "^region")]
region_map <- setNames(
  paste0("Region: ", str_replace_all(str_remove(region_terms, "^region"), "_", " ")),
  region_terms
)

weekend_term <- NULL
if ("isWeekend" %in% fixef_names) {
  weekend_term <- setNames(psqi_label_lookup["isWeekend"], "isWeekend")
} else if ("isWeekendTRUE" %in% fixef_names) {
  weekend_term <- setNames(psqi_label_lookup["isWeekend"], "isWeekendTRUE")
}

coef_map_psqi_global <- c(
  psqi_base_map,
  psqi_moderator_map,
  region_map,
  weekend_term,
  psqi_random_map
)
coef_map_psqi_global <- coef_map_psqi_global[!is.na(coef_map_psqi_global)]

psqi_models_list <- list(
  "Playtime (Imputed)" = model_psqi_global_imputed,
  "Chronotype × Playtime (Imputed)" = model_psqi_global_imputed_chrono,
  "Playtime (Complete-Case)" = model_psqi_global_completecase,
  "Chronotype × Playtime (Complete-Case)" = model_psqi_global_completecase_chrono
)

# Build custom rows: use Rubin bundles for imputed, single fits for complete-case
psqi_custom_rows <- {
  pooled_bundles <- list(
    "Playtime (Imputed)" = panel_psqi_rubin_base,
    "Chronotype × Playtime (Imputed)" = panel_psqi_rubin_chrono
  )
  cc_models <- list(
    "Playtime (Complete-Case)" = model_psqi_global_completecase,
    "Chronotype × Playtime (Complete-Case)" = model_psqi_global_completecase_chrono
  )
  pooled_rows <- get_custom_rows_pooled(pooled_bundles)
  cc_rows <- get_custom_rows(cc_models)
  # Merge: participant_df row then icc_df row
  merged <- pooled_rows
  for (col in names(cc_rows)) {
    if (col != "term") merged[[col]] <- cc_rows[[col]]
  }
  merged
}

psqi_group_cols <- seq_along(psqi_models_list) + 1

tt_psqi <- modelsummary(
  psqi_models_list,
  fmt = fmt_sprintf("%.2f"),
  estimate  = "{estimate} [{conf.low}, {conf.high}]{stars}",
  statistic = NULL,
  gof_omit = "AIC|BIC|Log|RMSE|R2|ICC|Num|nobs",
  coef_map = coef_map_psqi_global,
  add_rows = psqi_custom_rows,
  notes = c(
    "+ p < 0.10, * p < 0.05, ** p < 0.01, *** p < 0.001",
    "PSQI global score range: 0-21 (higher = worse sleep quality)",
    "Confidence intervals shown in brackets.",
    "ICC = Intraclass Correlation Coefficient (adjusted)."
  ),
  output = "tinytable"
) |>
  group_tt(
    j = list("PSQI Global" = psqi_group_cols),
    part = "header"
  )
dir.create("output/tables", showWarnings = FALSE, recursive = TRUE)
tryCatch(
  save_tt(tt_psqi, "output/tables/tbl-appendix-psqi-global.png", overwrite = TRUE),
  error = function(e) message("Note: PNG export skipped (", e$message, ")")
)
tt_psqi
```

```{=typst}
#pagebreak()
```

# Appendix: H2c Interaction Plot

```{r}
#| label: fig-h2c-interaction
#| fig-cap: "Predicted wellbeing (SWEMWBS) as a function of late-night gaming, moderated by chronotype. Lines show model-predicted SWEMWBS scores at three levels of chronotype: −1 SD (morning-oriented), mean, and +1 SD (evening-oriented). Ribbons indicate 95% confidence intervals. Predictions are derived from pooled H2c linear mixed model estimates (Rubin's rules across 20 imputed datasets), with other covariates held at their reference/mean values and random effects set to zero."
#| fig-width: 8
#| fig-height: 5
#| code-summary: "H2c interaction plot: Late-night gaming × Chronotype on Wellbeing"

library(lme4)
library(ggplot2)

# Use pooled coefficients and vcov from Rubin's rules
panel_h2c_fig_bundle <- if (!is.null(panel_h2c_bundle)) panel_h2c_bundle else
  tryCatch(readRDS("output/models/imputed_h2c_rubin.rds"), error = function(e) NULL)
panel_pv_h2c <- pool_vcov(panel_h2c_fig_bundle$fits)
fe <- panel_pv_h2c$qbar
vc <- panel_pv_h2c$vcov

# Chronotype levels: -1 SD, mean (0), +1 SD
# Use SD from first fit's data frame (identical across imputations)
chrono_sd <- sd(panel_h2c_fig_bundle$fits[[1]]@frame$msf_sc_centered)
chrono_levels <- c(-chrono_sd, 0, chrono_sd)
chrono_labels <- c("−1 SD (morning)", "Mean", "+1 SD (evening)")

# Late-night gaming range: 0 to 90 min (in 10-min units: 0 to 9)
ln_seq_10 <- seq(0, 9, by = 0.1)

# Build prediction grid
pred_grid <- expand.grid(
  ln_10 = ln_seq_10,
  chrono = chrono_levels
)
pred_grid$chrono_label <- factor(
  rep(chrono_labels, each = length(ln_seq_10)),
  levels = chrono_labels
)

# Predicted values: intercept + b_ln * x + b_chrono * c + b_int * x * c
# (all other covariates at 0/reference)
pred_grid$yhat <- fe["(Intercept)"] +
  fe["ln_biweekly_avg_minutes_played_10"] * pred_grid$ln_10 +
  fe["msf_sc_centered"] * pred_grid$chrono +
  fe["ln_biweekly_avg_minutes_played_10:msf_sc_centered"] * pred_grid$ln_10 * pred_grid$chrono

# Standard errors via delta method: var(yhat) = x' V x
# Design matrix columns: intercept, ln_10, chrono, ln_10:chrono
terms_idx <- c(
  "(Intercept)",
  "ln_biweekly_avg_minutes_played_10",
  "msf_sc_centered",
  "ln_biweekly_avg_minutes_played_10:msf_sc_centered"
)
V_sub <- vc[terms_idx, terms_idx]

pred_grid$se <- sapply(seq_len(nrow(pred_grid)), function(i) {
  x <- c(1, pred_grid$ln_10[i], pred_grid$chrono[i],
         pred_grid$ln_10[i] * pred_grid$chrono[i])
  sqrt(as.numeric(t(x) %*% V_sub %*% x))
})

pred_grid$ci_lo <- pred_grid$yhat - 1.96 * pred_grid$se
pred_grid$ci_hi <- pred_grid$yhat + 1.96 * pred_grid$se
pred_grid$ln_minutes <- pred_grid$ln_10 * 10

ggplot(pred_grid, aes(x = ln_minutes, y = yhat,
                      colour = chrono_label, fill = chrono_label)) +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), alpha = 0.15, colour = NA) +
  geom_line(linewidth = 1) +
  scale_x_continuous(
    breaks = seq(0, 90, 30),
    labels = paste0(seq(0, 90, 30), " min")
  ) +
  scale_colour_manual(
    values = c("−1 SD (morning)" = "#2166ac",
               "Mean" = "#4d4d4d",
               "+1 SD (evening)" = "#b2182b")
  ) +
  scale_fill_manual(
    values = c("−1 SD (morning)" = "#2166ac",
               "Mean" = "#4d4d4d",
               "+1 SD (evening)" = "#b2182b")
  ) +
  labs(
    x = "Average Daily Late-Night Gaming (min/day)",
    y = "Predicted SWEMWBS Score (7–35)",
    colour = "Chronotype",
    fill = "Chronotype",
    title = "Chronotype Moderates the Association Between\nLate-Night Gaming and Wellbeing (H2c)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

```{=typst}
#pagebreak()
```

# Appendix: Daily Diary Subsample

```{r}
#| label: tbl-diary-demographics
#| tbl-cap: "Daily Diary Subsample Characteristics"
#| echo: false
#| message: false
#| warning: false

library(tinytable)

# Define diary samples
diary_total_pids <- unique(diary_long$pid)
diary_analytical_pids <- unique(diary_ocat$pid)

# Person-level summaries for diary sample
diary_person_total <- diary_long |>
  group_by(pid) |>
  summarise(
    n_diary_entries = n(),
    ln_gaming_min_day = mean(late_night_gaming_hours_24h * 60, na.rm = TRUE),
    any_ln_gaming_pct = 100 * mean(late_night_gaming_hours_24h > 0, na.rm = TRUE),
    .groups = "drop"
  ) |>
  left_join(
    diary_long |> distinct(pid, age_scaled, bmi_scaled, SES_index_scaled, gender, region,
                            chronotype_hours, chronotype_centered, no_alarm_free),
    by = "pid"
  )

diary_person_analytical <- diary_ocat |>
  group_by(pid) |>
  summarise(
    n_diary_entries = n(),
    ln_gaming_min_day = mean(late_night_minutes_trim, na.rm = TRUE),
    any_ln_gaming_pct = 100 * mean(late_night_minutes_trim > 0, na.rm = TRUE),
    .groups = "drop"
  ) |>
  left_join(
    diary_ocat |> distinct(pid, age_scaled, bmi_scaled, SES_index_scaled, gender, region,
                            chronotype_hours, chronotype_centered, no_alarm_free),
    by = "pid"
  )

# Temporarily assign to master_data_total / master_data for helper functions
master_data_total_saved <- if (exists("master_data_total")) master_data_total else NULL
master_data_saved <- if (exists("master_data")) master_data else NULL
master_data_total <- diary_person_total
master_data <- diary_person_analytical

diary_summary_rows <- tibble(Characteristic = character(), Total = character(), Analytical = character())

# === A. Sociodemographics ===
diary_summary_rows <- bind_rows(diary_summary_rows,
  tibble(Characteristic = "**A. Sociodemographics**", Total = "", Analytical = ""))

diary_summary_rows <- bind_rows(diary_summary_rows,
  tibble(Characteristic = "N participants",
         Total = as.character(length(diary_total_pids)),
         Analytical = as.character(length(diary_analytical_pids))))

diary_summary_rows <- bind_rows(diary_summary_rows,
  tibble(Characteristic = "N diary entries",
         Total = as.character(nrow(diary_long)),
         Analytical = as.character(nrow(diary_ocat))))

diary_summary_rows <- bind_rows(diary_summary_rows,
  tibble(Characteristic = "    Diary entries per person (Mdn, IQR)",
         Total = sprintf("%.0f (%.0f)",
           median(diary_person_total$n_diary_entries),
           IQR(diary_person_total$n_diary_entries)),
         Analytical = sprintf("%.0f (%.0f)",
           median(diary_person_analytical$n_diary_entries),
           IQR(diary_person_analytical$n_diary_entries))))

# Use intake data for age (unscaled)
diary_intake_total <- data.intake |> filter(pid %in% diary_total_pids)
diary_intake_analytical <- data.intake |> filter(pid %in% diary_analytical_pids)

diary_summary_rows <- bind_rows(diary_summary_rows,
  tibble(Characteristic = "Age",
         Total = sprintf("%.1f (%.1f)",
           mean(as.numeric(diary_intake_total$age), na.rm = TRUE),
           sd(as.numeric(diary_intake_total$age), na.rm = TRUE)),
         Analytical = sprintf("%.1f (%.1f)",
           mean(as.numeric(diary_intake_analytical$age), na.rm = TRUE),
           sd(as.numeric(diary_intake_analytical$age), na.rm = TRUE))))

# Gender
diary_summary_rows <- bind_rows(diary_summary_rows,
  tibble(Characteristic = "Gender", Total = "", Analytical = ""))

diary_gender_counts_total <- diary_person_total |>
  count(gender, name = "n_tot") |> filter(!is.na(gender))
diary_gender_counts_ana <- diary_person_analytical |>
  count(gender, name = "n_ana") |> filter(!is.na(gender))
diary_gender_merged <- diary_gender_counts_total |>
  full_join(diary_gender_counts_ana, by = "gender") |>
  mutate(n_tot = replace_na(n_tot, 0), n_ana = replace_na(n_ana, 0))

for (g in diary_gender_merged$gender) {
  row <- diary_gender_merged |> filter(gender == g)
  diary_summary_rows <- bind_rows(diary_summary_rows,
    tibble(Characteristic = paste0("    ", g),
           Total = sprintf("%d (%.1f%%)", row$n_tot, 100 * row$n_tot / nrow(diary_person_total)),
           Analytical = sprintf("%d (%.1f%%)", row$n_ana, 100 * row$n_ana / nrow(diary_person_analytical))))
}

# BMI and SES
diary_summary_rows <- bind_rows(diary_summary_rows,
  add_continuous("bmi_scaled", "BMI (scaled)", 2))

diary_summary_rows <- bind_rows(diary_summary_rows,
  add_continuous("SES_index_scaled", "SES (scaled)", 2))

# === B. Chronotype ===
diary_summary_rows <- bind_rows(diary_summary_rows,
  tibble(Characteristic = "**B. Chronotype**", Total = "", Analytical = ""))

no_alarm_n_tot <- sum(master_data_total$no_alarm_free, na.rm = TRUE)
no_alarm_pct_tot <- 100 * no_alarm_n_tot / sum(!is.na(master_data_total$no_alarm_free))
no_alarm_n_ana <- sum(master_data$no_alarm_free, na.rm = TRUE)
no_alarm_pct_ana <- 100 * no_alarm_n_ana / sum(!is.na(master_data$no_alarm_free))
diary_summary_rows <- bind_rows(diary_summary_rows,
  tibble(Characteristic = "No alarm on free days",
         Total = sprintf("%d (%.1f%%)", no_alarm_n_tot, no_alarm_pct_tot),
         Analytical = sprintf("%d (%.1f%%)", no_alarm_n_ana, no_alarm_pct_ana)))

msf_med_tot <- median(master_data_total$chronotype_hours, na.rm = TRUE)
msf_iqr_tot <- IQR(master_data_total$chronotype_hours, na.rm = TRUE)
msf_med_ana <- median(master_data$chronotype_hours, na.rm = TRUE)
msf_iqr_ana <- IQR(master_data$chronotype_hours, na.rm = TRUE)
diary_summary_rows <- bind_rows(diary_summary_rows,
  tibble(Characteristic = "    MSF~sc~ (HH:MM)",
         Total = sprintf("%s (%s)", hours_to_hhmm(msf_med_tot), hours_to_hhmm(msf_iqr_tot)),
         Analytical = sprintf("%s (%s)", hours_to_hhmm(msf_med_ana), hours_to_hhmm(msf_iqr_ana))))

# === C. Gaming ===
diary_summary_rows <- bind_rows(diary_summary_rows,
  tibble(Characteristic = "**C. Gaming**", Total = "", Analytical = ""))

diary_summary_rows <- bind_rows(diary_summary_rows,
  add_median_iqr("ln_gaming_min_day", "LN gaming (min/day, Mdn, IQR)", 1))

diary_summary_rows <- bind_rows(diary_summary_rows,
  add_continuous("any_ln_gaming_pct", "% entries with any LN gaming", 1))

# === D. Outcomes ===
diary_summary_rows <- bind_rows(diary_summary_rows,
  tibble(Characteristic = "**D. Sleep Quality Distribution**", Total = "", Analytical = ""))

# Sleep quality distribution from diary entries (not person-level)
sq_dist_total <- diary_long |>
  filter(!is.na(sleep_quality)) |>
  count(sleep_quality, name = "n_tot")
sq_dist_ana <- diary_ocat |>
  mutate(sleep_quality = as.character(sleep_quality_ord)) |>
  count(sleep_quality, name = "n_ana")
sq_dist <- sq_dist_total |>
  full_join(sq_dist_ana, by = "sleep_quality") |>
  mutate(n_tot = replace_na(n_tot, 0), n_ana = replace_na(n_ana, 0))
sq_n_total <- sum(sq_dist$n_tot)
sq_n_ana <- sum(sq_dist$n_ana)
sq_order <- c("Very poor", "Poor", "Fair", "Good", "Very good")
sq_dist <- sq_dist |> arrange(match(sleep_quality, sq_order))

for (sq in sq_dist$sleep_quality) {
  row <- sq_dist |> filter(sleep_quality == sq)
  diary_summary_rows <- bind_rows(diary_summary_rows,
    tibble(Characteristic = paste0("    ", sq),
           Total = sprintf("%d (%.1f%%)", row$n_tot, 100 * row$n_tot / sq_n_total),
           Analytical = sprintf("%d (%.1f%%)", row$n_ana, 100 * row$n_ana / sq_n_ana)))
}

# Restore master_data_total / master_data
if (!is.null(master_data_total_saved)) master_data_total <- master_data_total_saved
if (!is.null(master_data_saved)) master_data <- master_data_saved

# Display table
diary_header_rows <- which(str_detect(diary_summary_rows$Characteristic, "^\\*\\*"))
diary_sub_item_rows <- which(str_detect(diary_summary_rows$Characteristic, "^    ") &
                             !str_detect(diary_summary_rows$Characteristic, "^        "))

diary_tbl <- diary_summary_rows |>
  tt(
    caption = "Daily Diary Subsample Characteristics",
    notes = "Values are M (SD) unless noted. LN = late-night. Sleep quality distribution counts are at the diary-entry level."
  ) |>
  format_tt(j = 1, markdown = TRUE) |>
  style_tt(j = 1, align = "l") |>
  style_tt(j = 2:3, align = "r") |>
  style_tt(i = diary_header_rows, bold = TRUE) |>
  format_tt(i = diary_header_rows, j = 1, fn = function(x) str_remove_all(x, "\\*\\*"))

if (length(diary_sub_item_rows) > 0) {
  diary_tbl <- diary_tbl |>
    style_tt(i = diary_sub_item_rows, j = 1, indent = 1) |>
    format_tt(i = diary_sub_item_rows, j = 1, fn = function(x) str_trim(x))
}

diary_tbl
```

