---
title: Late-Night Gaming, Sleep and Wellbeing
subtitle: Optional Subtitle
running-head: Short Title
author:
  - name: First Author
    email: author1@example.com
    orcid: 0000-0000-0000-0000
    url: https://example.com
    corresponding: true
    affiliation:
      - ref: 1
  - name: Second Author
    affiliation:
      - ref: 2
affiliations:
  - id: 1
    name: First University
    department: Department Name
  - id: 2
    name: Second University
abstract: |
  Your abstract goes here. This will appear prominently at the top of your document.
categories: [keyword1, keyword2, keyword3]
authornote: |
  Author notes go here.
date: last-modified
bibliography: bibliography.bib
format:
  html:
    toc: true
    toc-location: left
    theme: cosmo
    code-fold: true
    html-math-method: katex
    embed-resources: true
  pdf:
    documentclass: article
    toc: true
    number-sections: true
    colorlinks: true
  preprint-typst:
    wordcount: true
    citeproc: true
    # Common options - uncomment to use
    # theme: jou                   # Journal theme (2-column)
    # line-number: true            # Line numbering
    # fontsize: 11pt               # Font size
    # leading: 0.6em               # Line spacing
    # first-line-indent: 1.8em     # Paragraph indent
---

# Introduction

Concerns have been raised about the potential negative impacts of video gaming on sleep and overall wellbeing, particularly for adolescents and young adults and especially when gaming occurs late at night (Altintas et al., 2019; Exelmans & Van Den Bulck, 2015; Higuchi et al., 2005; King et al., 2013; Peracchia & Curcio, 2018). Late-night gaming has been shown to disrupt sleep patterns, reduce sleep duration, lower sleep quality, and increase daytime sleepiness (Exelmans & Van Den Bulck, 2015; Han et al., 2024; Kim, 2024; Kristensen et al., 2021).This is especially concerning given the far-reaching effects of sleep disturbances on cognitive and emotional functioning (Cain & Gradisar, 2010; LeBourgeois et al., 2017; McCoy & Strecker, 2011; Simon et al., 2020; Vriend et al., 2013). For instance, habitual gaming between 10 p.m. and 6 a.m. has been associated with an increased risk of depressive symptoms, partially mediated by daytime sleepiness (Lemola et al., 2011). Understanding the consequences of late-night gaming is thus vital for both gamers and health professionals.
Two key mechanisms have been proposed to explain the impact of late-night digital engagement—including gaming—on sleep. The first is the displacement hypothesis, which argues that late-night gaming is more harmful than daytime gaming because it cuts into sleep time (Twenge, 2019; Williams et al., 2008). Gamers often feel compelled to continue playing and struggle with self-regulation, which can lead to insufficient sleep (King & Delfabbro, 2009; Pirrone et al., 2024; Spada & Caselli, 2017). For example, adolescents experiencing a heightened sense of "flow" during challenging games delayed bedtime by up to 90 minutes  (Smith et al., 2017). 
The second mechanism involves arousal-related disturbances in sleep architecture caused by late-night gaming. Empirical studies have shown that extended gaming, especially when involving violent content, significantly decreases REM sleep and total sleep time (King et al., 2013). Weaver et al. (2010) highlighted that increased arousal levels due to pre-sleep gaming extend sleep latency and alter the natural progression into sleep stages. This delay in sleep onset could be exacerbated by lower melatonin levels following an evening of gaming, compared to neutral activities like board games, which are crucial for regulating the sleep-wake cycle (Hartmann et al., 2019).  
Negative effects of late-night gaming are often compounded among individuals with an eveningness chronotype—a group naturally predisposed to staying up late. Problematic gamers, who frequently possess this chronotype, are especially vulnerable to the detrimental effects of late-night gaming on sleep (Kristensen et al., 2021). Pre-sleep technology use may exacerbate the misalignment between their biological clock and societal demands by delaying sleep onset and reducing sleep duration, leading to poorer sleep quality and increased daytime sleepiness. Research has linked evening chronotype in adolescents to greater technology use at bedtime, in turn associated with delayed sleep onset, shorter sleep duration, and poorer sleep quality (Bruni et al., 2015; Gumport et al., 2021; Kortesoja et al., 2023; Reardon et al., 2023). Additionally, while Reardon et al. (2023) found that shorter sleep on weekdays was associated with greater psychological distress, technology medium and chronotype were not direct predictors of distress. Gumport et al. (2021) found that technology use improved emotional, social, cognitive, and physical health but worsened behavioral health, measured by the consumption of junk food, caffeine, alcohol, tobacco, and other substances, in evening-type adolescents. It remains unclear how strongly these findings apply to young adults and adults, as most research has focused on adolescent populations. This leaves an open question about the extent to which evening chronotypes in older age groups are similarly affected by pre-sleep technology use.
In sum, the literature indicates that video gaming, particularly when it occurs late at night, has significant implications for sleep quality, sleep duration, and overall wellbeing. This disruption can be attributed to both the displacement hypothesis (Twenge, 2019; Williams et al., 2008) and arousal-related disturbances in sleep architecture (King et al., 2013). Individual differences, such as chronotype, may moderate these effects, with eveningness chronotypes particularly vulnerable to the negative consequences of late-night gaming (Kristensen et al., 2021). The present study aims to empirically test the following hypotheses regarding the relationship between late-night gaming and sleep outcomes:

**H1**: Late-night gaming is associated with:

- **H1a**: Poorer sleep quality
- **H1b**: Shorter sleep duration
- **H1c**: Higher daytime sleepiness
- **H1d**: Lower wellbeing

In addition to testing direct relationships between late-night gaming and various sleep-related outcomes are critical to understand, we further assess the potential moderating role of chronotype, which refers to a person's natural preference for activities during certain times of the day—morningness or eveningness. Individuals with an evening chronotype tend to stay up later and may be more inclined to engage in late-night gaming, potentially exacerbating the negative impacts on sleep and wellbeing. The combination of an evening chronotype and late-night gaming may even have a compounded effect on overall wellbeing, as both factors are independently associated with poorer mental health outcomes. Given this, we propose the following:

**H2**: Chronotype moderates the relationships between late-night gaming and all outcomes in H1 (sleep quality, sleep duration, daytime sleepiness, and wellbeing), such that these negative associations are stronger for individuals with more of an eveningness chronotype.

By examining chronotype on a continuous scale as a moderating factor, this study seeks to provide a more nuanced understanding of the potential risks associated with late-night gaming and to identify individuals who may be most vulnerable to its negative effects.

# Methods

# Results

:::: {.place arguments='top, scope: "parent", float: true'}

```{r}
#| label: load-libraries

if (!require("pacman")) install.packages("pacman")
library(pacman)

p_load(tidyverse, lme4, marginaleffects, glmmTMB, mice, ordinal, modelsummary, lubridate, data.table, mctq)

studyStartDate <- as_datetime("2024-05-01 00:00:00")

# Source preprocessing functions
source("preprocess_data.R")

# TOGGLE HYPOTHESES EVALUATION
eval_h1 <- TRUE  # Set to FALSE to skip H1 hypotheses
eval_h2 <- FALSE  # Set to FALSE to skip H2 hypotheses

```

```{r}
#| label: load-data

# Load raw data
data.panel.raw <- read_csv("data/panel.csv.gz", show_col_types = FALSE)
data.intake.raw <- read_csv("data/intake.csv.gz", show_col_types = FALSE)

# Process intake data (demographics, scaled variables)
data.intake <- process_intake("data/intake.csv.gz")

# Process panel data (compute derived sleep/wellbeing variables, merge with intake)
data.panel <- process_panel("data/panel.csv.gz", data.intake) |>
  dplyr::select(pid, wave, region, psqi_6, starts_with(c("mctq", "eps", "psqi")),
                total_hours_sleep, wemwbs, epsTotal,
                age_scaled, bmi_scaled, SES_index_scaled, msf_sc_numeric, gender) |>
  mutate(
    psqi_6_ord = factor(psqi_6, ordered = TRUE),
    pid = as.character(pid)
  )

# Load gaming data
data.nin <- read_csv("data/nintendo.csv.gz", show_col_types = FALSE)
data.xbox <- read_csv("data/xbox.csv.gz", show_col_types = FALSE)
data.steam <- read_csv("data/steam.csv.gz", show_col_types = FALSE)

```

```{r}
#| label: merge data

# Standardize column names across gaming platforms
# Nintendo: session_start, session_end, duration
# Xbox: session_start, session_end, duration
# Steam: session_start, session_end (need to compute duration)

data.nin.std <- data.nin |>
  rename(sessionStart = session_start, sessionEnd = session_end) |>
  mutate(date = as_date(sessionStart))

data.xbox.std <- data.xbox |>
  rename(sessionStart = session_start, sessionEnd = session_end) |>
  mutate(date = as_date(sessionStart))

data.steam.std <- data.steam |>
  rename(sessionStart = session_start, sessionEnd = session_end) |>
  mutate(date = as_date(sessionStart))

# merge xbox, steam and nintendo
data.gaming <- bind_rows(data.xbox.std, data.steam.std, data.nin.std) |>

  mutate(

    # redefine day to begin and end at 4am, rather than midnight, so we can properly
    # assign late-night sessions to the previous calendar day
    dateRecoded = if_else(hour(sessionStart) < 6, date - 1, date),

    # calculate minutes_played for data.gaming using sessionEnd and sessionStart
    # create a binary variable for data.gaming called latenight if the sessionStart is between 23:00 and 06:00
    # create an isWeekend variable for data.gaming if the sessionStart is on a Saturday, Sunday or Friday
    minutes_played = as.numeric(difftime(sessionEnd, sessionStart, units = "mins")),
    latenight = ifelse(hour(sessionStart) >= 23 | hour(sessionStart) < 6, 1, 0),
    isWeekend = ifelse(weekdays(sessionStart) %in% c("Friday", "Saturday"), 1, 0),

    # calculate late night using lubridate::interval objects
    interval_gaming = interval(sessionStart, sessionEnd),
    interval_latenight = interval(dateRecoded + hours(23), dateRecoded + hours(30)), # gets the 11pm to 6am following day interval from date
    latenightMinutes = as.numeric(intersect(interval_gaming, interval_latenight))/60 # calculate overlap in minutes
  ) |>

  # assign each session to a wave based on the date, if the date is within 4
  # weeks of the start of the study, it is wave 2, if it is between 4 weeks and
  # 8 weeks it is wave 4, if it is between 8 weeks and 12 weeks it is wave 6
  mutate(
    days_since_start = as.numeric(difftime(sessionStart, studyStartDate, units = "days")),
    month = case_when(
      days_since_start >= 0 & days_since_start <= 28 ~ 2,
      days_since_start >= 29 & days_since_start <= 56 ~ 4,
      days_since_start >= 57 & days_since_start <= 84 ~ 6,
      TRUE ~ NA_real_
    ),
    # Define biweekly waves (14-day intervals) that end at day 84
    wave = case_when(
      days_since_start >= 0 & days_since_start <= 14 ~ 1,
      days_since_start >= 15 & days_since_start <= 28 ~ 2,
      days_since_start >= 29 & days_since_start <= 42 ~ 3,
      days_since_start >= 43 & days_since_start <= 56 ~ 4,
      days_since_start >= 57 & days_since_start <= 70 ~ 5,
      days_since_start >= 71 & days_since_start <= 84 ~ 6,
      TRUE ~ NA_real_  # Assign NA for days beyond 84
    ),
    pid = as.character(pid)
  ) 

# group gaming by PID, wave and latenight and calculate daily average minutes_played
gamingMonthly <- data.gaming |> 
  group_by(pid, month, isWeekend) %>%
  summarise(
    monthly_avg_minutes_played = sum(latenightMinutes)/28,
    monthly_total_minutes_played = sum(latenightMinutes),
  ) |> 
  left_join(data.panel |> dplyr::select(-msf_sc_numeric) |> filter(wave %in% c(2, 4, 6)), 
            by = c("pid","month" = "wave")) |> 
  left_join(data.panel |> dplyr::select(pid, msf_sc_numeric) |> filter(!is.na(msf_sc_numeric)), 
            by = c("pid")) |> # hacky way to make sure chronotype (measured in wave 1 only) doesn't get lost 
  mutate(
    monthly_avg_minutes_played = replace_na(monthly_avg_minutes_played, 0),
    monthly_total_minutes_played = replace_na(monthly_total_minutes_played, 0)
  ) |> 
  ungroup() |>
  arrange(as.integer(pid), month)

# group gaming by PID, wave and latenight and calculate daily average minutes_played
gamingBiweekly <- data.gaming |> 
  group_by(pid, wave, isWeekend) |> 
  summarise(
    biweekly_avg_minutes_played = sum(latenightMinutes)/14,
  ) |> 
  left_join(data.panel |> dplyr::select(-msf_sc_numeric), by = c("pid", "wave")) |> 
  left_join(data.panel |> dplyr::select(pid, msf_sc_numeric) |> filter(!is.na(msf_sc_numeric)), 
            by = c("pid")) |> # hacky way to make sure chronotype (measured in wave 1 only) doesn't get lost 
  mutate(biweekly_avg_minutes_played = replace_na(biweekly_avg_minutes_played, 0)) |> 
  ungroup() |> 
  arrange(as.integer(pid), wave) |> 
  dplyr::select(pid, wave, msf_sc_numeric, everything())

```

## Descriptive Raincloud Plots

Note: The dataset contains 7 WEMWBS items (SWEMWBS short form, range 7-35) rather than the full 14-item WEMWBS (range 14-70).

```{r}
#| label: raincloud-descriptives
#| code-summary: "Raincloud plots for key variables"
#| fig-width: 14
#| fig-height: 12

# Load additional packages for raincloud plots
if (!require("ggdist")) install.packages("ggdist")
library(ggdist)

# Prepare data for raincloud plots - select key variables
descriptive_data <- gamingMonthly |>
  dplyr::select(
    pid,
    month,
    monthly_avg_minutes_played,
    sleep_quality = psqi_6,
    sleep_duration = total_hours_sleep,
    daytime_sleepiness = epsTotal,
    wellbeing = wemwbs
  ) |>
  mutate(
    # Convert sleep_quality to numeric (1=Very good, 2=Fairly good, 3=Fairly bad, 4=Very bad)
    sleep_quality = as.numeric(factor(sleep_quality, 
                                      levels = c("Very good", "Fairly good", "Fairly bad", "Very bad"),
                                      ordered = TRUE)),
    # Label months for display
    month_label = factor(month, 
                        levels = c(2, 4, 6),
                        labels = c("Month 1", "Month 2", "Month 3"))
  ) |>
  pivot_longer(
    cols = c(monthly_avg_minutes_played, sleep_quality, sleep_duration, daytime_sleepiness, wellbeing),
    names_to = "variable",
    values_to = "value"
  ) |>
  filter(!is.na(value)) |>
  mutate(
    variable = factor(variable, 
                     levels = c("monthly_avg_minutes_played", "sleep_quality", 
                               "sleep_duration", "daytime_sleepiness", "wellbeing"),
                     labels = c("Late-Night Gaming\n(avg min/day)", 
                               "Sleep Quality\n(PSQI: 1-4)", 
                               "Sleep Duration\n(hours)",
                               "Daytime Sleepiness\n(EPS: 0-24)",
                               "Wellbeing\n(WEMWBS: 7-35)"))
  )

# Create a function for raincloud plot
create_raincloud <- function(data, var_name) {
  ggplot(data, aes(x = month_label, y = value, fill = month_label)) +
    # Half violin (density distribution)
    stat_halfeye(
      adjust = 0.5,
      width = 0.6,
      .width = 0,
      justification = -0.2,
      point_colour = NA,
      alpha = 0.7
    ) +
    # Boxplot
    geom_boxplot(
      width = 0.15,
      outlier.shape = NA,
      alpha = 0.5
    ) +
    # Styling
    scale_fill_brewer(palette = "Set2", name = "Month") +
    labs(
      title = var_name,
      y = "Value",
      x = "Monthly Wave"
    ) +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "bottom",
      panel.grid.minor = element_blank(),
      plot.title = element_text(face = "bold", size = 13),
      axis.text.x = element_text(size = 10)
    )
}

# Create individual plots for each variable
plot_gaming <- descriptive_data |> 
  filter(variable == "Late-Night Gaming\n(avg min/day)") |>
  create_raincloud("Late-Night Gaming (avg min/day)")

plot_sleep_quality <- descriptive_data |> 
  filter(variable == "Sleep Quality\n(PSQI: 1-4)") |>
  create_raincloud("Sleep Quality (PSQI: 1-4)")

plot_sleep_duration <- descriptive_data |> 
  filter(variable == "Sleep Duration\n(hours)") |>
  create_raincloud("Sleep Duration (hours)")

plot_sleepiness <- descriptive_data |> 
  filter(variable == "Daytime Sleepiness\n(EPS: 0-24)") |>
  create_raincloud("Daytime Sleepiness (EPS: 0-24)")

plot_wellbeing <- descriptive_data |> 
  filter(variable == "Wellbeing\n(WEMWBS: 7-35)") |>
  create_raincloud("Wellbeing (WEMWBS: 7-35)")

# Display plots
plot_gaming
plot_sleep_quality
plot_sleep_duration
plot_sleepiness
plot_wellbeing

# Print summary statistics by month
descriptive_data |>
  group_by(variable, month_label) |>
  summarise(
    N = n(),
    Mean = mean(value, na.rm = TRUE),
    SD = sd(value, na.rm = TRUE),
    Median = median(value, na.rm = TRUE),
    Min = min(value, na.rm = TRUE),
    Max = max(value, na.rm = TRUE),
    .groups = "drop"
  ) |>
  knitr::kable(digits = 2, caption = "Descriptive Statistics for Key Variables by Monthly Wave")
```


## H1a: Late-night gaming is associated with poorer sleep quality.

Multilevel ordinal regression whereby monthly average minutes played predicts sleep quality (PSQI), controlling for age, BMI, SES index, region, and whether playtime falls on a weekend, with a random intercept and slope for participants.

```{r}
#| label: h1a
#| code-summary: "Fit H1a model"
#| eval: !expr eval_h1

# For psqi_6 1 means Very good, 2 means Fairly good, 3 means Fairly bad, 4 means Very bad

# Fit the model with rescaled covariates
model.h1a <- clmm(psqi_6_ord ~ monthly_avg_minutes_played + (1 + monthly_avg_minutes_played | pid) +
                    age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend,
                  data = gamingMonthly)

modelsummary(
  list(`Model H1a` = model.h1a),
  fmt = 2,
  estimate  = "{estimate} [{conf.low}, {conf.high}]{stars}", 
  statistic = NULL
)
```


## H1b: Late-night gaming is associated with shorter sleep duration.

Multilevel linear regression whereby monthly average minutes played predicts total hours of sleep (PSQI), controlling for age, BMI, SES index, region, gender, and whether playtime falls on a weekend, with a random intercept and slope for participants and a random intercept for gender.

```{r}
#| label: h1b
#| code-summary: "Fit H1b model"

# Fit the model
model.h1b <- lmer(total_hours_sleep ~ monthly_avg_minutes_played + (1 + monthly_avg_minutes_played | pid) +
                    age_scaled + bmi_scaled + SES_index_scaled + (1 | gender) + region + isWeekend, 
                  data = gamingMonthly
                  )

# Summarize the model using modelsummary
modelsummary(
  list(`Model H1b` = model.h1b),
  fmt = 2,
  estimate  = "{estimate} [{conf.low}, {conf.high}]{stars}", 
  statistic = NULL
)
```

## H1c: Late-night gaming is associated with lower well-being.

Multilevel linear regression whereby monthly average minutes played predicts daytime sleepiness (Epworth Sleepiness Scale), controlling for age, BMI, SES index, region, gender, and whether playtime falls on a weekend, with a random intercept and slope for participants and a random intercept for gender.

```{r}
#| label: h1c
#| code-summary: "Fit H1c model"
#| eval: !expr eval_h1

model.h1c <- lmer(wemwbs ~ biweekly_avg_minutes_played + (1 + biweekly_avg_minutes_played | pid) +
                    age_scaled + bmi_scaled + SES_index_scaled + (1 | gender) + region + isWeekend, 
                  data = gamingBiweekly)

modelsummary(
  list(`Model H1c` = model.h1c),
  fmt = 2,
  estimate  = "{estimate} [{conf.low}, {conf.high}]{stars}", 
  statistic = NULL
)

```

## H1d: Late-night gaming is associated with higher daytime sleepiness.

Multilevel linear regression whereby biweekly average minutes played predicts well-being (WEMWBS), controlling for age, BMI, SES index, region, gender, and whether playtime falls on a weekend, with a random intercept and slope for participants and a random intercept for gender.

```{r}
#| label: h1d
#| code-summary: "Fit H1d model"
#| eval: !expr eval_h1

# Fit the model
model.h1d <- lmer(epsTotal ~ monthly_avg_minutes_played + (1 + monthly_avg_minutes_played | pid) +
                    age_scaled + bmi_scaled + SES_index_scaled + (1 | gender) + region + isWeekend, 
                  data = gamingMonthly)

# Summarize the model using modelsummary
modelsummary(
  list(`Model H1d` = model.h1d),
  fmt = 2,
  estimate  = "{estimate} [{conf.low}, {conf.high}]{stars}", 
  statistic = NULL
)
```

## H2

### H2a: The negative association between late-night gaming and sleep quality is more pronounced among evening chronotypes.

Multilevel ordinal regression whereby the interaction between monthly average minutes played and chronotype (MSFsc: mid-sleep on free days corrected for sleep debt on weekdays; MCTQ) predicts sleep quality (PSQI), controlling for age, BMI, SES index, region, and whether playtime falls on a weekend, with a random intercept and slope for participants.

On the simulation data, there are convergence problems, so here we center late night minutes and chronotype. This will either be uncentered for interpretation or not necessary on the true data.)

```{r}
#| label: h2a
#| code-summary: "Fit H2a model"
#| eval: !expr eval_h2

# Fit the model with rescaled covariates
model.h2a <- clmm(psqi_6_ord ~ scale(monthly_avg_minutes_played) * scale(msf_sc_numeric) + (1 + monthly_avg_minutes_played | pid) +
                    age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend,
                  data = gamingMonthly)

modelsummary(
  list(`Model H2a` = model.h2a),
  fmt = 2,
  estimate  = "{estimate} [{conf.low}, {conf.high}]{stars}", 
  statistic = NULL
)
```

## H2b: The negative association between late-night gaming and sleep duration is more pronounced among evening chronotypes. 

```{r}
#| label: h2b
#| code-summary: "Fit H2b model"
#| eval: !expr eval_h2

# Fit the model
model.h2b <- lmer(total_hours_sleep ~ monthly_avg_minutes_played * msf_sc_numeric + (1 + monthly_avg_minutes_played | pid) +
                    age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend + (1 | gender), 
                  data = gamingMonthly)

# Summarize the model using modelsummary
modelsummary(
  list(`Model H2b` = model.h2b),
  fmt = 2,
  estimate  = "{estimate} [{conf.low}, {conf.high}]{stars}", 
  statistic = NULL
)

```

### H2c: The negative association between late-night gaming and well-being is more pronounced among evening chronotypes.

Multilevel linear regression whereby the interaction between monthly average minutes played and chronotype (MSFsc; MCTQ) predicts daytime sleepiness (Epworth Sleepiness Scale), controlling for age, BMI, SES index, region, gender, and whether playtime falls on a weekend, with a random intercept and slope for participants and a random intercept for gender.

```{r}
#| label: h2c
#| code-summary: "Fit H2c model"
#| eval: !expr eval_h2

# Fit the model
model.h2c <- lmer(wemwbs ~ biweekly_avg_minutes_played * msf_sc_numeric + (1 | pid) +
                    age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend + (1 | gender), 
                  data = gamingBiweekly)
modelsummary(
  list(`Model h2c` = model.h2c),
  fmt = 2,
  estimate  = "{estimate} [{conf.low}, {conf.high}]{stars}", 
  statistic = NULL
)

```

### H2d: The negative association between late-night gaming and daytime sleepiness is more pronounced among evening chronotypes.

Multilevel linear regression whereby the interaction between biweekly average minutes played and chronotype (MSFsc; MCTQ) predicts well-being (WEMWBS), controlling for age, BMI, SES index, region, gender, and whether playtime falls on a weekend, with a random intercept for participants and a random intercept for gender.

```{r}
#| label: h2d
#| code-summary: "Fit H2d model"
#| eval: !expr eval_h2

model.h2d <- lmer(epsTotal ~ monthly_avg_minutes_played * msf_sc_numeric + (1 + monthly_avg_minutes_played | pid) +
                    age_scaled + bmi_scaled + SES_index_scaled + region + isWeekend + (1 | gender), 
                  data = gamingMonthly)
                  
modelsummary(
  list(`Model H2d` = model.h2d),
  fmt = 2,
  estimate  = "{estimate} [{conf.low}, {conf.high}]{stars}", 
  statistic = NULL
)
```


## Precision Analysis


In the above examples, we use the simulated data, which mirrors the structure of the true data, but lacks control over the distribution and relationships between the particular variables used in this study. To provide rough indications of the estimated precision of our tests, here we simulate a dataset with a known relationship between playtime and sleep, and then fit H1b as above (selecting one hypothesis for detailed inspection to illustrate).

We assume that late-night gaming is zero-inflated gamma distributed, where 30% of people will have no late-night play at all, and the remainder generally very little (0-30 minutes), with a handful of extreme values. We assume a true effect of 1-hour of late-night gaming reducing sleep duration by .2 hours, and that the effect varies by person, with a standard deviation of .2. We then fit the model as above, and examine the estimated effect size and confidence intervals.

```{r}
#| label: precision-analysis
#| code-summary: "Show the code (precision analysis)"

b_sleep <- -.2
sd_sleep <- 1
sd_play <- .5
p_noplay <- .3
sleep_sd <- 1
random_intercept_sd <- .6
random_slope_sd <- .2

data_sim <- gamingMonthly |> 
  dplyr::select(pid, isWeekend, total_hours_sleep, age_scaled, bmi_scaled, SES_index_scaled, gender, region) |>
  group_by(pid) |>
  mutate(
    random_intercept = rnorm(n(), mean = 0, sd = random_intercept_sd), 
    random_slope = rnorm(n(), mean = 0, sd = random_slope_sd)
  ) |> 
  ungroup() |> 
  mutate(
    monthly_avg_hours_played = ifelse(rbinom(n(), 1, p_noplay) == 1, 0, rgamma(n(), shape = 2, scale = .3)),
    sleep_effect = (b_sleep + random_slope) * monthly_avg_hours_played,
    total_hours_sleep_sim = rnorm(n(), mean = 7.5 + sleep_effect + random_intercept, sd = sleep_sd),
    total_hours_sleep_sim = ifelse(is.na(total_hours_sleep), NA, total_hours_sleep_sim) # restore missingness
  )

model.h1b <- lmer(total_hours_sleep_sim ~ monthly_avg_hours_played + (1 + monthly_avg_hours_played | pid) +
                    age_scaled + bmi_scaled + SES_index_scaled + (1 | gender) + region + isWeekend, 
                  data = data_sim
                  )

modelsummary(
  list(`Model H1b` = model.h1b),
  fmt = 2,
  estimate  = "{estimate} [{conf.low}, {conf.high}]{stars}", 
  statistic = NULL
)
plot_predictions(model.h1b, condition = "monthly_avg_hours_played")
```

::::

# Discussion

# References

:::{#refs}
:::

```{=typst}
#show: appendix.with()
```
# Appendix content here
